---
title: "cDiag4"
author: "Daniel Zoleikhaeian"
date: "2023-02-10"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up environment

```{r setting up environment}
library(ggplot2)
library(dplyr)
library(plyr)

setwd('C:/Users/danie/Documents/Joshi Lab Materials/3 Studies Dataset/Dataset Merge/012623')
df <- read.csv('bgc_merge_cDiag123_new.csv')
#View(head(df))
df_LA_gpc <- df[(df$cLocationCity == 'Los Angeles') & (df$cStudy == 'GPC'), ]

#View(head(df_LA_gpc))
```

## Overall counts in LA

```{r}
library(readxl)
df_check_2010 <- read_xlsx('reproj_30_2010.xlsx')
df_check_2020 <- read_xlsx('reproj_30_2020.xlsx')

LA_2010 <- df_check_2010[which(df_check_2010$City == 'Los Angeles, CA'),]
LA_2020 <- df_check_2020[which(df_check_2020$City == 'Los Angeles, CA'),]
LA <- rbind(LA_2010, LA_2020)
LA <- LA[,-1]
#View(LA)

LA_avg <- as.data.frame(sapply(LA, mean, margin = 2))
#View(LA_avg)
LA_avg$categ <- rownames(LA_avg)
#View(LA_avg)
rownames(LA_avg) <- NULL
freq <- LA_avg[,1]
LA_avg[,1] <- LA_avg[,2]
LA_avg[,2] <- freq

colnames(LA_avg) <- c('categ', 'freq')
#View(LA_avg)

# Standardizing race acronyms
LA_avg_races <- LA_avg[1:7,]
colnames(LA_avg_races) <- c('cRace', 'freq')
LA_avg_races$cRace[LA_avg_races$cRace == 'NA'] <- 'AE'
LA_avg_races$cRace[LA_avg_races$cRace == 'PI'] <- 'NH'
LA_avg_races <- LA_avg_races[order(LA_avg_races$cRace), ]
rownames(LA_avg_races) <- NULL
LA_avg_races <- LA_avg_races[1:6, ]

# Race Proportions
LA_avg_races$Proportions <- LA_avg_races$freq / sum(LA_avg_races$freq)

LA_avg_races$cLocationInstitution <- 'Average Census'

LA_avg_races <- LA_avg_races[,c(1,4,2,3)]
colnames(LA_avg_races)[3] <- 'total_count'

# #View(LA_avg_races)

# Ethnicity
LA_avg_eth <- LA_avg[8:9, ]
colnames(LA_avg_eth) <- c('cHispanicorLatino', 'freq')
LA_avg_eth$cHispanicorLatino[LA_avg_eth$cHispanicorLatino == 'NH'] <- 'No'
LA_avg_eth$cHispanicorLatino[LA_avg_eth$cHispanicorLatino == 'HA'] <- 'Yes'

LA_avg_eth$Proportions <- LA_avg_eth$freq / sum(LA_avg_eth$freq)

LA_avg_eth$cLocationInstitution <- 'Average Census'

LA_avg_eth <- LA_avg_eth[,c(1,4,2,3)]
colnames(LA_avg_eth)[3] <- 'total_count'
#View(LA_avg_eth)

# Gender
LA_avg_gen <- LA_avg[10:11,]
colnames(LA_avg_gen) <- c('cGender', 'freq')

LA_avg_gen$Proportions <- LA_avg_gen$freq / sum(LA_avg_gen$freq)

LA_avg_gen$cLocationInstitution <- 'Average Census'

LA_avg_gen <- LA_avg_gen[,c(1,4,2,3)]
colnames(LA_avg_gen)[3] <- 'total_count'
#View(LA_avg_gen)
```


## Counting by race and diagnosis4 within each institution
```{r}
brdI <- df_LA_gpc %>% 
  dplyr::group_by(cRace,cDiagnosis4, cLocationInstitution) %>%  
  dplyr::summarise(total_count=n(), .groups = 'drop')

View(brdI)
```

### Examining SZSAFD across race and institution

```{r}
szsafd <- brdI[brdI$cDiagnosis4 == 'SZSAFD', ]
#View(szsafd)
szsafd_colors <- c("#808080", "#85AD00", "#00A5FF", "#F8766D")

inst <- levels(factor(szsafd$cLocationInstitution))
szsafd_totals <- rep(0, length(inst))

for (i in 1:length(inst)) {
  szsafd_totals[i] <- sum(szsafd$total_count[szsafd$cLocationInstitution == inst[i]])
}

names(szsafd_totals) <- inst
szsafd_totals

props <- rep(0, nrow(szsafd))

for (i in 1:nrow(szsafd)) {
  props[i] <- szsafd$total_count[i] / szsafd_totals[names(szsafd_totals) == szsafd$cLocationInstitution[i]]
}

szsafd$Proportions <- props
szsafd$cDiagnosis4 <- NULL

#View(szsafd)

szsafd_tot <- rbind(szsafd, LA_avg_races)
#View(szsafd_tot)

szsafd_tots <- ggplot(data = szsafd_tot, aes(x = cRace, y = Proportions, fill = cLocationInstitution)) +
  geom_bar(position = position_dodge(), 
           stat = 'identity') +
  scale_fill_manual(values = szsafd_colors) +
  ggtitle('Proportions of Race for SZSAFD') + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

szsafd_tots
```

### BAD12 across race and instution 

```{r}
bad12 <- brdI[brdI$cDiagnosis4 == 'BAD12', ]
#View(bad12)

inst <- levels(factor(bad12$cLocationInstitution))
bad12_totals <- rep(0, length(inst))

for (i in 1:length(inst)) {
  bad12_totals[i] <- sum(bad12$total_count[bad12$cLocationInstitution == inst[i]])
}

names(bad12_totals) <- inst
bad12_totals

props <- rep(0, nrow(bad12))

for (i in 1:nrow(bad12)) {
  props[i] <- bad12$total_count[i] / bad12_totals[names(bad12_totals) == bad12$cLocationInstitution[i]]
}

bad12$Proportions <- props
bad12$cDiagnosis4 <- NULL

#View(bad12)

bad12_tot <- rbind(bad12, LA_avg_races)
#View(bad12_tot)

bad12_tots <- ggplot(data = bad12_tot, aes(x = cRace, y = Proportions, fill = cLocationInstitution)) +
  geom_bar(position = position_dodge(), 
           stat = 'identity') +
  scale_fill_manual(values = szsafd_colors) +
  ggtitle('Proportions of Race for BAD12') + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

bad12_tots
```

### CS across race and institution

```{r}
cs <- brdI[brdI$cDiagnosis4 == 'CS', ]
cs_colors <- c("#808080", "#00A5FF", "#F8766D")
#View(cs)

inst <- levels(factor(cs$cLocationInstitution))
cs_totals <- rep(0, length(inst))

for (i in 1:length(inst)) {
  cs_totals[i] <- sum(cs$total_count[cs$cLocationInstitution == inst[i]])
}

names(cs_totals) <- inst
cs_totals

props <- rep(0, nrow(cs))

for (i in 1:nrow(cs)) {
  props[i] <- cs$total_count[i] / cs_totals[names(cs_totals) == cs$cLocationInstitution[i]]
}

cs$Proportions <- props
cs$cDiagnosis4 <- NULL

#View(cs)

cs_tot <- rbind(cs, LA_avg_races)
#View(cs_tot)

cs_tots <- ggplot(data = cs_tot, aes(x = cRace, y = Proportions, fill = cLocationInstitution)) +
  geom_bar(position = position_dodge(), 
           stat = 'identity') +
  scale_fill_manual(values = cs_colors) +
  ggtitle('Proportions of Race in Controls') + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

cs_tots
```



## Counting by gender and cDiagnosis4 across instutions

```{r}
bgdI <- df_LA_gpc %>% 
  dplyr::group_by(cGender, cDiagnosis4, cLocationInstitution) %>% 
  dplyr::summarise(total_count=n(), .groups = 'drop')

#View(bgdI)
```

### Examining SZSAFD

```{r}
szsafd <- bgdI[bgdI$cDiagnosis4 == 'SZSAFD', ]
#View(szsafd)

inst <- levels(factor(szsafd$cLocationInstitution))
szsafd_totals <- rep(0, length(inst))

for (i in 1:length(inst)) {
  szsafd_totals[i] <- sum(szsafd$total_count[szsafd$cLocationInstitution == inst[i]])
}

names(szsafd_totals) <- inst
szsafd_totals

props <- rep(0, nrow(szsafd))

for (i in 1:nrow(szsafd)) {
  props[i] <- szsafd$total_count[i] / szsafd_totals[names(szsafd_totals) == szsafd$cLocationInstitution[i]]
}

szsafd$Proportions <- props
szsafd$cDiagnosis4 <- NULL

szsafd_tot <- rbind(szsafd, LA_avg_gen)

#View(szsafd)

szsafd_tots <- ggplot(data = szsafd_tot, aes(x = cGender, y = Proportions, fill = cLocationInstitution)) +
  geom_bar(position = position_dodge(), 
           stat = 'identity') +
  scale_fill_manual(values = szsafd_colors) +
  ggtitle('Proportions of Gender for SZSAFD') + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

szsafd_tots
```

### Examining BAD12

```{r}
bad12 <- bgdI[bgdI$cDiagnosis4 == 'BAD12', ]
#View(bad12)

inst <- levels(factor(bad12$cLocationInstitution))
bad12_totals <- rep(0, length(inst))

for (i in 1:length(inst)) {
  bad12_totals[i] <- sum(bad12$total_count[bad12$cLocationInstitution == inst[i]])
}

names(bad12_totals) <- inst
bad12_totals

props <- rep(0, nrow(bad12))

for (i in 1:nrow(bad12)) {
  props[i] <- bad12$total_count[i] / bad12_totals[names(bad12_totals) == bad12$cLocationInstitution[i]]
}

bad12$Proportions <- props
bad12$cDiagnosis4 <- NULL

#View(bad12)

bad12_tot <- rbind(bad12, LA_avg_gen)

bad12_tots <- ggplot(data = bad12_tot, aes(x = cGender, y = Proportions, fill = cLocationInstitution)) +
  geom_bar(position = position_dodge(), 
           stat = 'identity') +
  scale_fill_manual(values = szsafd_colors) +
  ggtitle('Proportions of Gender for BAD12') + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

bad12_tots
```

### Examining CS
```{r}
cs <- bgdI[bgdI$cDiagnosis4 == 'CS', ]
#View(cs)

inst <- levels(factor(cs$cLocationInstitution))
cs_totals <- rep(0, length(inst))

for (i in 1:length(inst)) {
  cs_totals[i] <- sum(cs$total_count[cs$cLocationInstitution == inst[i]])
}

names(cs_totals) <- inst
cs_totals

props <- rep(0, nrow(cs))

for (i in 1:nrow(cs)) {
  props[i] <- cs$total_count[i] / cs_totals[names(cs_totals) == cs$cLocationInstitution[i]]
}

cs$Proportions <- props
cs$cDiagnosis4 <- NULL

#View(cs)

cs_tot <- rbind(cs, LA_avg_gen)

cs_tots <- ggplot(data = cs_tot, aes(x = cGender, y = Proportions, fill = cLocationInstitution)) +
  geom_bar(position = position_dodge(), 
           stat = 'identity') +
  scale_fill_manual(values = cs_colors) +
  ggtitle('Proportions of Gender for Controls') + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

cs_tots
```




## Counting by ethnicity and cDiagnosis4 across instutions
```{r}
bedI <- df_LA_gpc %>% 
  dplyr::group_by(cHispanicorLatino, cDiagnosis4, cLocationInstitution) %>% 
  dplyr::summarise(total_count=n(), .groups = 'drop')

#View(bedI)
```

### Examining SZSAFD
```{r}
szsafd <- bedI[bedI$cDiagnosis4 == 'SZSAFD', ]
#View(szsafd)

inst <- levels(factor(szsafd$cLocationInstitution))
szsafd_totals <- rep(0, length(inst))

for (i in 1:length(inst)) {
  szsafd_totals[i] <- sum(szsafd$total_count[szsafd$cLocationInstitution == inst[i]])
}

names(szsafd_totals) <- inst
szsafd_totals

props <- rep(0, nrow(szsafd))

for (i in 1:nrow(szsafd)) {
  props[i] <- szsafd$total_count[i] / szsafd_totals[names(szsafd_totals) == szsafd$cLocationInstitution[i]]
}

szsafd$Proportions <- props
szsafd$cDiagnosis4 <- NULL

#View(szsafd)
szsafd_tot <- rbind(szsafd, LA_avg_eth)

szsafd_tots <- ggplot(data = szsafd_tot, aes(x = cHispanicorLatino, y = Proportions, fill = cLocationInstitution)) +
  geom_bar(position = position_dodge(), 
           stat = 'identity') +
  scale_fill_manual(values = szsafd_colors) +
  ggtitle('Proportions of Ethnicity for SZSAFD') + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

szsafd_tots
```

### Examining BAD12
```{r}
bad12 <- bedI[bedI$cDiagnosis4 == 'BAD12', ]
#View(bad12)

inst <- levels(factor(bad12$cLocationInstitution))
bad12_totals <- rep(0, length(inst))

for (i in 1:length(inst)) {
  bad12_totals[i] <- sum(bad12$total_count[bad12$cLocationInstitution == inst[i]])
}

names(bad12_totals) <- inst
bad12_totals

props <- rep(0, nrow(bad12))

for (i in 1:nrow(bad12)) {
  props[i] <- bad12$total_count[i] / bad12_totals[names(bad12_totals) == bad12$cLocationInstitution[i]]
}

bad12$Proportions <- props
bad12$cDiagnosis4 <- NULL

#View(bad12)
bad12_tot <- rbind(bad12, LA_avg_eth)

bad12_tots <- ggplot(data = bad12_tot, aes(x = cHispanicorLatino, y = Proportions, fill = cLocationInstitution)) +
  geom_bar(position = position_dodge(), 
           stat = 'identity') +
  scale_fill_manual(values = szsafd_colors) +
  ggtitle('Proportions of Ethnicity for BAD12') + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

bad12_tots
```

### Examining CS

```{r}
cs <- bedI[bedI$cDiagnosis4 == 'CS', ]
#View(cs)

inst <- levels(factor(cs$cLocationInstitution))
cs_totals <- rep(0, length(inst))

for (i in 1:length(inst)) {
  cs_totals[i] <- sum(cs$total_count[cs$cLocationInstitution == inst[i]])
}

names(cs_totals) <- inst
cs_totals

props <- rep(0, nrow(cs))

for (i in 1:nrow(cs)) {
  props[i] <- cs$total_count[i] / cs_totals[names(cs_totals) == cs$cLocationInstitution[i]]
}

cs$Proportions <- props
cs$cDiagnosis4 <- NULL

#View(cs)
cs_tot <- rbind(cs, LA_avg_eth)

cs_tots <- ggplot(data = cs_tot, aes(x = cHispanicorLatino, y = Proportions, fill = cLocationInstitution)) +
  geom_bar(position = position_dodge(), 
           stat = 'identity') +
  scale_fill_manual(values = cs_colors) +
  ggtitle('Proportions of Ethnicity for Controls') + 
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

cs_tots
```

