---
title: "cDiag3_UCLA"
author: "Daniel Zoleikhaeian"
date: "2023-02-02"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up environment

```{r setting up environment}
library(ggplot2)
library(dplyr)
library(plyr)

df <- read.csv('bgc_merge_cDiag123_new.csv')

df_LA <- df[df$cLocationCity == 'Los Angeles', ]
levels(factor(df_LA$cLocationInstitution))
df_La_counts <- as.data.frame(plyr::count(df_LA, 'cLocationInstitution'))
df_La_diags <- as.data.frame(plyr::count(df_LA, 'cDiagnosis3'))
View(df_La_diags)

df_LA_bdI <- df_LA %>% 
  dplyr::group_by(cLocationInstitution, cDiagnosis3) %>%
  dplyr::summarise(total_count=n(), .groups = 'drop')

View(df_LA_bdI)

View(df_La_counts)

df_gpc <- df[df$cStudy == 'GPC', ]
df_ucla <- df[df$cLocationInstitution == 'UCLA', ]
df_ucla_gpc <- df_ucla[df_ucla$cStudy == 'GPC', ]
df_ucla_cogs2 <- df_ucla[df_ucla$cStudy == 'COGS2', ]
```

## Counting by race and diagnosis

```{r cDiag3 by race and diagnosis for gpc and cogs2}

# counting elements by race and diagnosis
gpc_cDiag3_ucla <- df_ucla_gpc %>% 
  dplyr::group_by(cRace, cDiagnosis3) %>%  
  dplyr::summarise(total_count=n(), .groups = 'drop')

cogs2_cDiag3_ucla <- df_ucla_cogs2 %>% 
  dplyr::group_by(cRace, cDiagnosis3) %>%  
  dplyr::summarise(total_count=n(), .groups = 'drop')

# generating proportions
gpc_cDiag3_ucla$proportion <- gpc_cDiag3_ucla$total_count / sum(gpc_cDiag3_ucla$total_count)
cogs2_cDiag3_ucla$proportion <- cogs2_cDiag3_ucla$total_count / sum(cogs2_cDiag3_ucla$total_count)

# View(gpc_cDiag3_ucla)
# View(cogs2_cDiag3_ucla)
```

## Differences in UCLA CS between groups 

```{r differences in controls}

cs_gpc <- gpc_cDiag3_ucla[gpc_cDiag3_ucla$cDiagnosis3 == 'CS',]
cs_cogs2 <- cogs2_cDiag3_ucla[cogs2_cDiag3_ucla$cDiagnosis3 == 'CS',]

cs_gpc$cStudy <- rep('GPC', nrow(cs_gpc))
cs_cogs2$cStudy <- rep('COGS2', nrow(cs_cogs2))

# View(cs_gpc)
# View(cs_cogs2)
cs_ucla <- rbind(cs_gpc, cs_cogs2)

race_counts_cs <- ggplot(data = cs_ucla, aes(x = cRace, y = total_count, fill = cStudy)) +
  geom_bar(position = position_dodge(), 
           stat = 'identity') +
  scale_fill_manual(values = c('orange', 'darkgreen')) +
  geom_text(aes(label = total_count), vjust = 0) + 
  ggtitle('Counts of Race in Controls') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# race_counts_cs
```

## Examining cDiagnosis3

### Examining cDiagnosis3 Counts

```{r plotting counts}

df_combined <- rbind(df_ucla_gpc, df_ucla_cogs2)

df_combined <- df_combined %>% 
  dplyr::group_by(cStudy, cDiagnosis3) %>%  
  dplyr::summarise(total_count=n(), .groups = 'drop')

df_combined$Proportion

overall_cDiag3 <- ggplot(data = df_combined, aes(x = cDiagnosis3, y = total_count, fill = cStudy)) +
  geom_bar(position = position_dodge(), 
           stat = 'identity') +
  scale_fill_manual(values = c('orange', 'darkgreen')) +
  geom_text(aes(label = total_count), vjust = 0) + 
  ggtitle('Counts of cDiagnosis3') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# overall_cDiag3
```

### Examining cDiagnosis3 Proportions

```{r plotting proportions}
gpc_la_diag3 <- as.data.frame(plyr::count(df_ucla_gpc, 'cDiagnosis3'))
cogs2_la_diag3 <- as.data.frame(plyr::count(df_ucla_cogs2, 'cDiagnosis3'))

gpc_la_diag3$Proportion <- gpc_la_diag3$freq / sum(gpc_la_diag3$freq)
gpc_la_diag3$cStudy <- rep('GPC', nrow(gpc_la_diag3))

cogs2_la_diag3$Proportion <- cogs2_la_diag3$freq / sum(cogs2_la_diag3$freq)
cogs2_la_diag3$cStudy <- rep('COGS2', nrow(cogs2_la_diag3))

props_combined <- rbind(gpc_la_diag3, cogs2_la_diag3)

props_cDiag3 <- ggplot(data = props_combined, aes(x = cDiagnosis3, y = Proportion, fill = cStudy)) +
  geom_bar(position = position_dodge(), 
           stat = 'identity') +
  scale_fill_manual(values = c('orange', 'darkgreen')) +
  ggtitle('Proportions of cDiagnosis3') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))

# props_cDiag3
```

## Examining Race and cDiagnosis3 Interaction between Studies

```{r by race, study, and cDiagnosis3}
b_rcd <- df_ucla %>% 
  dplyr::group_by(cRace, cDiagnosis3, cStudy) %>%  
  dplyr::summarise(total_count=n(), .groups = 'drop')
# View(b_rcd)
```

```{r in African Americans}
# African Americans
b_aa <- b_rcd[b_rcd$cRace == 'AA', ]
b_aa <- b_aa[order(b_aa$cStudy), ]
b_aa$Proportion <- c(b_aa$total_count[b_aa$cStudy == 'COGS2'] / sum(b_aa$total_count[b_aa$cStudy == 'COGS2']), b_aa$total_count[b_aa$cStudy == 'GPC'] / sum(b_aa$total_count[b_aa$cStudy == 'GPC']))
# View(b_aa)

aa_plot <- ggplot(data = b_aa, aes(x = cDiagnosis3, y = Proportion, fill = cStudy)) +
  geom_bar(position = position_dodge(), 
           stat = 'identity') +
  scale_fill_manual(values = c('orange', 'darkgreen')) +
  ggtitle('Proportions of cDiagnosis3 in African Americans') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))  

# aa_plot
```

```{r  in American Indians/Alaska Natives}
b_ae <- b_rcd[b_rcd$cRace == 'AE', ]
b_ae <- b_ae[order(b_ae$cStudy), ]
b_ae$Proportion <- c(b_ae$total_count[b_ae$cStudy == 'COGS2'] / sum(b_ae$total_count[b_ae$cStudy == 'COGS2']), b_ae$total_count[b_ae$cStudy == 'GPC'] / sum(b_ae$total_count[b_ae$cStudy == 'GPC']))
# View(b_ae)

ae_plot <- ggplot(data = b_ae, aes(x = cDiagnosis3, y = Proportion, fill = cStudy)) +
  geom_bar(position = position_dodge(), 
           stat = 'identity') +
  scale_fill_manual(values = c('orange', 'darkgreen')) +
  ggtitle('Proportions of cDiagnosis3 in American Indians/Alaska Natives') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))  
# all had SZSAFD
```

```{r Asian Americans}
b_as <- b_rcd[b_rcd$cRace == 'AS', ]

b_as <- b_as[order(b_as$cStudy), ]
b_as$Proportion <- c(b_as$total_count[b_as$cStudy == 'COGS2'] / sum(b_as$total_count[b_as$cStudy == 'COGS2']), b_as$total_count[b_as$cStudy == 'GPC'] / sum(b_as$total_count[b_as$cStudy == 'GPC']))
# View(b_as)

as_plot <- ggplot(data = b_as, aes(x = cDiagnosis3, y = Proportion, fill = cStudy)) +
  geom_bar(position = position_dodge(), 
           stat = 'identity') +
  scale_fill_manual(values = c('orange', 'darkgreen')) +
  ggtitle('Proportions of cDiagnosis3 in Asian Americans') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))  

# as_plot
```

```{r in Caucasians}
b_ca <- b_rcd[b_rcd$cRace == 'CA', ]
b_ca <- b_ca[order(b_ca$cStudy), ]
b_ca$Proportion <- c(b_ca$total_count[b_ca$cStudy == 'COGS2'] / sum(b_ca$total_count[b_ca$cStudy == 'COGS2']), b_ca$total_count[b_ca$cStudy == 'GPC'] / sum(b_ca$total_count[b_ca$cStudy == 'GPC']))
# View(b_ca)

ca_plot <- ggplot(data = b_ca, aes(x = cDiagnosis3, y = Proportion, fill = cStudy)) +
  geom_bar(position = position_dodge(), 
           stat = 'identity') +
  scale_fill_manual(values = c('orange', 'darkgreen')) +
  ggtitle('Proportions of cDiagnosis3 in Caucasian Americans') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))  

# ca_plot
```

```{r Multiracial}
b_mr <- b_rcd[b_rcd$cRace == 'MR', ]
b_mr <- b_mr[order(b_mr$cStudy), ]
b_mr$Proportion <- c(b_mr$total_count[b_mr$cStudy == 'COGS2'] / sum(b_mr$total_count[b_mr$cStudy == 'COGS2']), b_mr$total_count[b_mr$cStudy == 'GPC'] / sum(b_mr$total_count[b_mr$cStudy == 'GPC']))
# View(b_mr)

mr_plot <- ggplot(data = b_mr, aes(x = cDiagnosis3, y = Proportion, fill = cStudy)) +
  geom_bar(position = position_dodge(), 
           stat = 'identity') +
  scale_fill_manual(values = c('orange', 'darkgreen')) +
  ggtitle('Proportions of cDiagnosis3 in Multiracial Americans') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))  

# mr_plot
```

```{r in Native Hawaiians/Pacific Islanders}
b_nh <- b_rcd[b_rcd$cRace == 'NH', ]
b_nh <- b_nh[order(b_nh$cStudy), ]
b_nh$Proportion <- c(b_nh$total_count[b_nh$cStudy == 'COGS2'] / sum(b_nh$total_count[b_nh$cStudy == 'COGS2']), b_nh$total_count[b_nh$cStudy == 'GPC'] / sum(b_nh$total_count[b_nh$cStudy == 'GPC']))
# View(b_nh)

nh_plot <- ggplot(data = b_nh, aes(x = cDiagnosis3, y = Proportion, fill = cStudy)) +
  geom_bar(position = position_dodge(), 
           stat = 'identity') +
  scale_fill_manual(values = c('orange', 'darkgreen')) +
  ggtitle('Proportions of cDiagnosis3 in Native Hawaiians/Pacific Islanders') +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))  

# nh_plot

```

## Plot Collection

```{r Plots}
race_counts_cs
overall_cDiag3
props_cDiag3
aa_plot
ae_plot
as_plot
ca_plot
mr_plot
nh_plot
```

## Main Findings for UCLA

* GPC has a higher proportion of SZSAFD than COGS2
* GPC lacks healthy controls for many races
* COGS2 has more balanced proportions of controls to patients across all races
* COGS2 has more Caucasians in controls 