---
title: "Dif_idx_across_size"
author: "Daniel Zoleikhaeian"
date: "2023-02-24"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading libraries
```{r}
library(plyr)
library(dplyr)
library(readxl)
library(ggplot2)
```

## Working directory
```{r}
setwd("C:/Users/danie/Documents/Joshi Lab Materials/3 Studies Dataset/Dataset Merge/012623")
```

## Loading Datasets and creating average dataset
```{r}
total_df <- read.csv('bgc_merge_cDiag123_new.csv')

# Changing La Jolla to San Diego
total_df$cLocationCity[total_df$cLocationCity == 'La Jolla'] <- 'San Diego'
# Changing Brooklyn to New York
total_df$cLocationCity[total_df$cLocationCity == 'Brooklyn'] <- 'New York'

# Removing USC/MASS
total_df <- total_df[total_df$cLocationCity != 'USC/MASS', ]

df_check_2010 <- read_xlsx('reproj_30_2010.xlsx')
df_check_2020 <- read_xlsx('reproj_30_2020.xlsx')

dem_2010 <- df_check_2010[,-1]
dem_2020 <- df_check_2020[,-1]

df_avg <- df_check_2010
df_avg[,] <- NA
df_avg$City <- NULL

for (i in 1:ncol(dem_2010)) {
  df_avg[,i] <- rowMeans(cbind(dem_2010[,i], dem_2020[,i]))
  
}

df_avg <- cbind(df_check_2010$City, df_avg)
colnames(df_avg)[1] <- 'City'


# Reformatting cities
for (i in 1:length(df_avg$City)) {
  df_avg$City[i] <- substr(df_avg$City[i], 1, unlist(gregexpr(',', df_avg$City[i]))[1] - 1)
}

df_avg$City[df_avg$City == 'Stonybrook'] <- 'Stony Brook'

# Standardizing race acronyms
colnames(df_avg)[colnames(df_avg) == 'NA'] <- 'AE'
colnames(df_avg)[colnames(df_avg) == 'NH'] <- 'Not Hispanic'
colnames(df_avg)[colnames(df_avg) == 'PI'] <- 'NH'

#View(df_avg)

# Re-run with only GPC and COGS 

# Looking only at SZSAFD
szsafd <- total_df[total_df$cDiagnosis4 == 'SZSAFD', ]
cs <- total_df[total_df$cDiagnosis4 == 'CS', ]
bad12 <- total_df[total_df$cDiagnosis4 == 'BAD12',]
mdd <- total_df[total_df$cDiagnosis4 == 'MDD', ]


# Looking at city intersection between dataset and 
test_cities <- intersect(df_avg$City, szsafd$cLocationCity)
test_cities
```

## Dependencies: Analysis methodology - simulation for later use
```{r}
# Modified input: df_null is now a vector

msim <- function(df_obs, df_null, races) {
  null_races <- colnames(df_null)[2:8]
  null_dist <- rep(null_races, round(df_null[1,2:8],0)) # edited this line
  results <- matrix(0,6,1000)

  for (i in 1:1000) {
    sam <- sample(null_dist, nrow(df_obs), replace = F)
    race_sam <- as.data.frame(plyr::count(sam))
    
    not_represented <- c()
  
    for (race in races) {
      if (!is.element(race, race_sam$x))
        not_represented <- c(not_represented, race)
      else
        next
    }
    
    for (race in not_represented) {
      race_sam <- rbind(race_sam, data.frame(x = race, freq = 0))
    }
    
    race_sam <- race_sam[order(race_sam$x),]
    rownames(race_sam) <- race_sam$x
    race_sam$x <- NULL
    
    results[,i] <- race_sam$freq
    results[,i] <- results[,i] / sum(results[,i])
  }

  sigs <- rep(NA, 6)
  # two-tailed test with p-value = 0.05/6 = 0.008
  # generate a 99.2% CI
  # see if the observed value lies outside of this interval
  lower_bounds <- c()
  upper_bounds <- c()
  
  for (i in 1:nrow(df_obs)) {
    row_res <- sort(results[i, ])
    lower_bound <- row_res[4]
    lower_bounds <- c(lower_bounds, lower_bound)
    upper_bound <- row_res[996]
    upper_bounds <- c(upper_bounds, upper_bound)
    
    
    if (df_obs$prop[i] > upper_bound || df_obs$prop[i] < lower_bound) {
      sigs[i] <- T
    }
    else {
      sigs[i] <- F
    }
  }
  
  df_obs$lower <- lower_bounds
  df_obs$upper <- upper_bounds
  df_obs$Significance <- sigs
  return(df_obs)
}

```

## Starting analysis: prop dif across city for SZSAFD_gpc

```{r}
szsafd_gpc <- szsafd[szsafd$cStudy == 'GPC', ]
test_cities <- intersect(df_avg$City, szsafd_gpc$cLocationCity)

null_race_cts <- df_avg[, 1:8]
null_race_props <- null_race_cts

for (i in 1:nrow(null_race_cts)) {
  city_tot <- sum(null_race_cts[i, 2:8])
  null_race_props[i, 2:8] <- null_race_cts[i,2:8] / city_tot
}

#null_race_props <- as.data.frame(apply(null_race_cts[,-1], MARGIN = 2, function(x) {x / sum(x)}))


View(null_race_props)
df_city <- data.frame(City = null_race_cts$City)

null_race_props <- null_race_props[null_race_props$City %in% test_cities,]
null_race_props <- null_race_props[order(null_race_props$City), ]


brc <- szsafd_gpc %>%  
  dplyr::group_by(cRace, cLocationCity) %>%  
  dplyr::summarise(total_count=n(), .groups = 'drop')

brc <- brc[brc$cLocationCity %in% test_cities, ]
 

brc <- brc[order(brc$cLocationCity), ]
brc$Prop <- NA

for (i in 1:length(test_cities)) {
    brc$Prop[brc$cLocationCity == test_cities[i]] <- brc$total_count[brc$cLocationCity == test_cities[i]] / sum(brc$total_count[brc$cLocationCity == test_cities[i]])
}

# View(brc)
# View(null_race_props)


```

### AA SZSAFD GPC (tester code)
```{r}
AA_row_null <- as.data.frame(null_race_props[,c('City','AA')])
AA_obs <- brc[brc$cRace == 'AA', ]

View(AA_row_null)

# Verifying positional matching among cities
(levels(factor(AA_obs$cLocationCity)) == AA_row_null$City)

#View(AA_obs)
AA_row_obs <- AA_obs$Prop

propdif_AA <- AA_row_obs - AA_row_null$AA

city_cts <- plyr::count(szsafd$cLocationCity)

ct_to_plot <- city_cts[city_cts$x %in% test_cities, ]
#View(city_cts)

# Verifying positional matching
ct_to_plot$x == AA_row_null$City

plot(log(ct_to_plot$freq), propdif_AA, main = 'SZSAFD: Proportion Difference by Sample Size for AA', xlab = 'Log(City Sample Size)', ylab = 'Proportion Difference (Observed - Census Average)' )

AA_lr <- lm(propdif_AA ~ log(ct_to_plot$freq))
summary(AA_lr)
abline(AA_lr, col = 'red', lwd = 2)

mtext(paste('gpc Adj R-squared=', round(summary(AA_lr)$adj.r.squared,3)), side = 3)

plot(AA_lr, 1)
plot(AA_lr, 4)

## Alternate analysis: use the city population of AA
View(null_race_cts)

null_race_cts <- null_race_cts[null_race_cts$City %in% test_cities, ]
AA_null <- null_race_cts$AA

plot(log(AA_null), propdif_AA, main = 'SZSAFD: Proportion Difference by AA City Population for AA', xlab = 'log(AA population)', ylab = 'Proportion Difference (Observed - Census Average)' )

AA_total_pop_lr <- lm(propdif_AA ~ log(AA_null))
summary(AA_total_pop_lr)
abline(AA_total_pop_lr, col = 'red')
mtext(paste('gpc Adj R-squared=', round(summary(AA_total_pop_lr)$adj.r.squared,3)), side = 3)

## Alternate analysis: use total city population instead
null_race_sums <- as.data.frame(apply(null_race_cts[,-1], 1, sum))
colnames(null_race_sums) <- 'Total City Population'
View(null_race_sums)

Total_pop_lr_AA <- lm(propdif_AA ~ log(null_race_sums[,1]))
summary(Total_pop_lr_AA)

plot(log(null_race_sums[,1]), propdif_AA, main = 'SZSAFD: Proportion Difference by Total City Population for AA', xlab = 'Total City Population', ylab = 'Proportion Difference (Observed - Census Average)' )
text(log(null_race_sums[,1]), propdif_AA, 1:nrow(null_race_sums), pos=1)

abline(Total_pop_lr_AA, col = 'red')
mtext(paste('gpc Adj R-squared=', round(summary(Total_pop_lr_AA)$adj.r.squared,3)), side = 3)

```

## Function for getting plot and lm data
```{r}

race_cities_plot_data <- function(null_df, obs_df, race_acr) {
  
  obs_cities <- obs_df$cLocationCity[obs_df$cRace == race_acr]
  
  test_cities <- intersect(null_df$City, obs_cities)
  
  null_race_cts <- null_df[, 1:8]
  null_race_props <- null_race_cts

  #null_race_props <- as.data.frame(apply(null_race_cts[,-1], MARGIN = 2, function(x) {x / sum(x)}))
  
  #df_city <- data.frame(City = null_race_cts$City)
  
  #null_race_props <- cbind(df_city, null_race_props)
  for (i in 1:nrow(null_race_cts)) {
  city_tot <- sum(null_race_cts[i, 2:8])
  null_race_props[i, 2:8] <- null_race_cts[i,2:8] / city_tot
}
  
  null_race_props <- null_race_props[null_race_props$City %in% test_cities,]
  null_race_props <- null_race_props[order(null_race_props$City), ]
  
  brc <- obs_df %>%  
    dplyr::group_by(cRace, cLocationCity) %>%  
    dplyr::summarise(total_count=n(), .groups = 'drop')
  
  brc <- brc[brc$cLocationCity %in% test_cities, ]
  
  brc <- brc[order(brc$cLocationCity), ]
  brc$Prop <- NA
  
  # Turning counts into proportions
  for (i in 1:length(test_cities)) {
    brc$Prop[brc$cLocationCity == test_cities[i]] <- brc$total_count[brc$cLocationCity == test_cities[i]] / sum(brc$total_count[brc$cLocationCity == test_cities[i]])
  }
  
  null_row <- as.data.frame(null_race_props[,c('City', race_acr)])
  obs_races <- brc[brc$cRace == race_acr, ]
  obs_prop <- obs_races$Prop
  
  ## Begin output variables to send to lm() and plot()
  propdif <- obs_prop - null_row[,2] ## response
  city_cts <- plyr::count(obs_df$cLocationCity)
  ct_to_plot <- city_cts$freq[city_cts$x %in% test_cities] ## predictor 1
  
  null_race_cts <- null_race_cts[null_race_cts$City %in% test_cities, ]
  race_of_interest_null <- null_race_cts[, colnames(null_race_cts) == race_acr] ## predictor 2
  
  null_race_sums <- as.data.frame(apply(null_race_cts[,-1], 1, sum)) ## predictor 3
  colnames(null_race_sums) <- 'Total City Population'
  
  output <- list(propdif, ct_to_plot, race_of_interest_null, null_race_sums[,1])
  names(output) <- c('Prop dif', 'city sample', 'census average race count', 'census average pop total')
  
  return (output)
  
}

plot_race_lm <- function(data_list, race_acr, disease, study) {
  plot(log(data_list[[2]]), data_list[[1]], main = paste(disease, ': Proportion Difference by Sample Size for ', race_acr, sep = ''), xlab = 'Log(City Sample Size)', ylab = 'Proportion Difference (Observed - Census Average)' )
    text(log(data_list[[2]]), data_list[[1]], 1:length(data_list[[2]]), pos=4, offset = 0.25)

  
  city_sam_lr <- lm(data_list[[1]] ~ log(data_list[[2]]))
  abline(city_sam_lr, col = 'red', lwd = 2)
  

  mtext(paste(study, 'Adj R-squared=', round(summary(city_sam_lr)$adj.r.squared,3)), side = 3)
  plot(city_sam_lr, 1)
  plot(city_sam_lr, 4)
  
  plot(log(data_list[[3]]), data_list[[1]], main = paste(disease, ': Proportion Difference by City Population for ', race_acr, sep = ''), xlab = paste('log(',race_acr,' population)', sep = ''), ylab = 'Proportion Difference (Observed - Census Average)' )
  text(log(data_list[[3]]), data_list[[1]], 1:length(data_list[[3]]), pos=4, offset = 0.25)


  race_total_pop_lr <- lm(data_list[[1]] ~ log(data_list[[3]]))
  abline(race_total_pop_lr, col = 'red')
  mtext(paste(study, 'Adj R-squared=', round(summary(race_total_pop_lr)$adj.r.squared,3)), side = 3)
  plot(race_total_pop_lr, 1)
  plot(race_total_pop_lr, 4)
  
  Total_city_pop <- lm(data_list[[1]] ~ log(data_list[[4]]))

  plot(log(data_list[[4]]), data_list[[1]], main = paste(disease, ': Proportion Difference in ', race_acr, ' by Total City Population', sep = ''), xlab = 'Total City Population', ylab = 'Proportion Difference (Observed - Census Average)' )
    text(log(data_list[[4]]), data_list[[1]], 1:length(data_list[[4]]), pos=4, offset = 0.25)

  
  abline(Total_city_pop, col = 'red')
  mtext(paste(study, 'Adj R-squared=', round(summary(Total_city_pop)$adj.r.squared,3)), side = 3)
  plot(Total_city_pop, 1)
  plot(Total_city_pop, 4)
  
  return(list(city_sam_lr, race_total_pop_lr, Total_city_pop))
  
}

```

## Examples applying the custom functions
```{r AA Analysis}
szsafd_gpc <- szsafd[szsafd$cStudy == 'GPC', ]

AA_out <- race_cities_plot_data(df_avg, szsafd_gpc, 'AA')
plot_race_lm(AA_out,'AA', 'SZSAFD', 'GPC')

```

```{r}
AS_out <- race_cities_plot_data(df_avg, szsafd_gpc, 'AS')
lm_res_AS <- plot_race_lm(AS_out,'AS', 'SZSAFD', 'GPC')

```

```{r}
CA_out <- race_cities_plot_data(df_avg, szsafd_gpc, 'CA')
plot_race_lm(CA_out,'CA', 'SZSAFD', 'GPC')
```


## Starting Analysis: prop dif across city for SZSAFD
```{r city counts across total df}
city_cts <- plyr::count(szsafd$cLocationCity)
city_cts

miss_cit <- setdiff(df_avg$City, city_cts$x)
add_tocity <- data.frame(x = miss_cit,
                         freq = rep(0, length(miss_cit)))
city_cts <- rbind(city_cts, add_tocity)
city_cts <- city_cts[city_cts$x != setdiff(city_cts$x, df_avg$City),]

city_cts <- city_cts[order(city_cts$x), ]

# Extracting only races from the null dataframe

null_race_cts <- df_avg[, 1:8]
#View(null_race_cts)
rownames(null_race_cts) <- null_race_cts$City
null_race_cts$City <- NULL

null_race_cts <- t(null_race_cts)

null_race_props <- as.data.frame(apply(null_race_cts, MARGIN = 2, function(x) {x / sum(x)}))
szsafd_cogs2 <- szsafd[szsafd$cStudy == 'COGS2', ]
szsafd_bsnip <- szsafd[szsafd$cStudy == 'BSNIP1' | szsafd$cStudy =='BSNIP2', ]
szsafd_gpc <- szsafd[szsafd$cStudy == 'GPC', ]
# View(null_race_props)

missing_cities <- setdiff(df_avg$City, szsafd_gpc$cLocationCity)
zero_cts <- data.frame(cRace = rep(c('AA', 'AS', 'CA'),length(missing_cities)),
                       cLocationCity = rep(missing_cities, 3),
                       total_count = rep(0,3 * length(missing_cities)))

# count up the subjects in each city by race
brc <- szsafd_gpc %>%  
  dplyr::group_by(cRace, cLocationCity) %>%  
  dplyr::summarise(total_count=n(), .groups = 'drop')

brc <- rbind(brc, zero_cts)
#brc <- brc[brc$cLocationCity != setdiff(brc$cLocationCity, df_avg$City),]

brc <- brc[order(brc$cLocationCity), ]

brc$Prop <- NA
View(brc)


# turn into proportions
for (i in 1:length(test_cities)) {
 if (sum(brc$total_count[brc$cLocationCity ==  test_cities2[i]]) != 0) {
    brc$Prop[brc$cLocationCity == test_cities2[i]] <- brc$total_count[brc$cLocationCity == test_cities2[i]] / sum(brc$total_count[brc$cLocationCity == test_cities2[i]])
  }
  else {
    brc$Prop[brc$cLocationCity == test_cities2[i]] <- 0
  }
  
}
```

### AA SZSAFD
```{r}
AA_row_null <- as.data.frame(t(null_race_props['AA',]))
#View(AA_row_null)
AA_obs <- brc[brc$cRace == 'AA', ]
AA_obs <- na.omit(AA_obs)

# Verifying positional matching among cities
sum(levels(factor(AA_obs$cLocationCity)) == rownames(AA_row_null))


#View(AA_obs)
AA_row_obs <- AA_obs$Prop

propdif_AA <- AA_row_obs - AA_row_null$AA


ct_to_plot <- city_cts[city_cts$x %in% test_cities, ]
#View(city_cts)

# Verifying positional matching
ct_to_plot$x == rownames(AA_row_null)

plot(log(ct_to_plot$freq), propdif_AA, main = 'SZSAFD: Proportion Difference by Sample Size for AA', xlab = 'Log(City Sample Size)', ylab = 'Proportion Difference (Observed - Census Average)' )

AA_lr <- lm(propdif_AA ~ log(ct_to_plot$freq))
summary(AA_lr)
abline(AA_lr, col = 'red', lwd = 2)

mtext(paste('gpc Adj R-squared=', round(summary(AA_lr)$adj.r.squared,3)), side = 3)

plot(AA_lr, 1)
plot(AA_lr, 4)

```

### AS SZSAFD
```{r}
AS_row_null <- as.data.frame(t(null_race_props['AS',]))
#View(AS_row_null)
AS_obs <- brc[brc$cRace == 'AS', ]
AS_obs <- na.omit(AS_obs)

# Verifying positional matching among cities
sum(levels(factor(AS_obs$cLocationCity)) == rownames(AS_row_null))


#View(AS_obs)
AS_row_obs <- AS_obs$Prop

propdif_AS <- AS_row_obs - AS_row_null$AS


ct_to_plot <- city_cts[city_cts$x %in% test_cities, ]
#View(city_cts)

# Verifying positional matching
ct_to_plot$x == rownames(AS_row_null)

plot(log(ct_to_plot$freq), propdif_AS, main = 'SZSAFD: Proportion Difference by Sample Size for AS', xlab = 'Log(City Sample Size)', ylab = 'Proportion Difference (Observed - Census Average)' )

AS_lr <- lm(propdif_AS ~ log(ct_to_plot$freq))
summary(AS_lr)
abline(AS_lr, col = 'red', lwd = 2)

mtext(paste('Adj R-squared=', round(summary(AS_lr)$adj.r.squared,3)), side = 3)

plot(AS_lr, 1)
plot(AS_lr, 4)

```

### CA SZSAFD
```{r}
CA_row_null <- as.data.frame(t(null_race_props['CA',]))
#View(CA_row_null)
CA_obs <- brc[brc$cRace == 'CA', ]
CA_obs <- na.omit(CA_obs)

# Verifying positional matching among cities
sum(levels(factor(CA_obs$cLocationCity)) == rownames(CA_row_null))


#View(CA_obs)
CA_row_obs <- CA_obs$Prop

propdif_CA <- CA_row_obs - CA_row_null$CA

ct_to_plot <- city_cts[city_cts$x %in% test_cities, ]
#View(city_cts)

# Verifying positional matching
ct_to_plot$x == rownames(CA_row_null)

plot(log(ct_to_plot$freq), propdif_CA, main = 'SZSAFD: Proportion Difference by Sample Size for CA', xlab = 'Log(City Sample Size)', ylab = 'Proportion Difference (Observed - Census Average)' )

CA_lr <- lm(propdif_CA ~ log(ct_to_plot$freq))
summary(CA_lr)
abline(CA_lr, col = 'red', lwd = 2)

mtext(paste('Adj R-squared=', round(summary(CA_lr)$adj.r.squared,3)), side = 3)

plot(CA_lr, 1)
plot(CA_lr, 4)

```

## Repeated analysis for CS
```{r}

# Looking at city intersection between dataset and 
test_cities <- intersect(df_avg$City, cs$cLocationCity)
test_cities

# Checking which cities are missing
missing_cities <- setdiff(df_avg$City, cs$cLocationCity)
zero_cts <- data.frame(cRace = rep(c('AA', 'AS', 'CA'),2),
                       cLocationCity = rep(missing_cities, 3),
                       total_count = rep(0,6))
# View(zero_cts)

## Augusta and Baltimore don't have CS
```

### Begin total counts
```{r city counts across total df}
city_cts <- plyr::count(cs$cLocationCity)
miss_cit <- setdiff(df_avg$City, city_cts$x)
add_tocity <- data.frame(x = miss_cit,
                         freq = rep(0, length(miss_cit)))
city_cts <- rbind(city_cts, add_tocity)
city_cts <- city_cts[city_cts$x != setdiff(city_cts$x, df_avg$City),]

city_cts <- city_cts[order(city_cts$x), ]
rownames(city_cts) <- NULL
View(city_cts)

# Extracting only races from the null dataframe

null_race_cts <- df_avg[, 1:8]
#View(null_race_cts)
rownames(null_race_cts) <- null_race_cts$City
null_race_cts$City <- NULL

null_race_cts <- t(null_race_cts)


null_race_props <- as.data.frame(apply(null_race_cts, MARGIN = 2, function(x) {x / sum(x)}))

# View(null_race_props)

# count up the subjects in each city by race
brc <- cs %>%  
  dplyr::group_by(cRace, cLocationCity) %>%  
  dplyr::summarise(total_count=n(), .groups = 'drop')

brc <- rbind(brc, zero_cts)
brc <- brc[brc$cLocationCity != setdiff(brc$cLocationCity, df_avg$City),]

brc <- brc[order(brc$cLocationCity), ]

brc$Prop <- NA
#View(brc)


# turn into proportions
for (i in 1:length(test_cities2)) {
  if (sum(brc$total_count[brc$cLocationCity ==  test_cities2[i]]) != 0) {
    brc$Prop[brc$cLocationCity == test_cities2[i]] <- brc$total_count[brc$cLocationCity == test_cities2[i]] / sum(brc$total_count[brc$cLocationCity == test_cities2[i]])
  }
  else {
    brc$Prop[brc$cLocationCity == test_cities2[i]] <- 0
  }
  
}
```

### AA CS
```{r}
AA_row_null <- as.data.frame(t(null_race_props['AA',]))
#View(AA_row_null)

# still_miss <- setdiff(rownames(AA_row_null), AA_obs$cLocationCity)
# add_zero <- data.frame(cRace = rep('AA',length(still_miss)),
#                        cLocationCity = still_miss,
#                        total_count = rep(0,length(still_miss)),
#                        Prop = rep(0, length(still_miss)))
# 
# brc <- rbind(brc, add_zero)

AA_obs <- brc[brc$cRace == 'AA', ]

# get rid of cities 
AA_obs <- na.omit(AA_obs)

View(AA_obs)

AA_obs <- AA_obs[order(AA_obs$cLocationCity), ]

# Verifying positional matching among cities
levels(factor(AA_obs$cLocationCity)) == rownames(AA_row_null)

setdiff(AA_obs$cLocationCity, rownames(AA_row_null))
#View(AA_obs)
#View(AA_row_null)
AA_row_obs <- AA_obs$Prop

propdif_AA <- AA_row_obs - AA_row_null$AA


ct_to_plot <- city_cts[city_cts$x %in% test_cities2, ]
#View(city_cts)

# Verifying positional matching
ct_to_plot$x == rownames(AA_row_null)

plot(sqrt(ct_to_plot$freq), propdif_AA, main = 'CS: Proportion Difference by Sample Size for AA', xlab = 'sqrt(City Sample Size)', ylab = 'Proportion Difference (Observed - Census Average)' )

AA_lr <- lm(propdif_AA ~ sqrt(ct_to_plot$freq))
summary(AA_lr)
abline(AA_lr, col = 'red', lwd = 2)

mtext(paste('Adj R-squared=', round(summary(AA_lr)$adj.r.squared,3)), side = 3)

plot(AA_lr, 1)
plot(AA_lr, 4)

```

### AS CS
```{r}
AS_row_null <- as.data.frame(t(null_race_props['AS',]))

still_miss <- setdiff(rownames(AS_row_null), AS_obs$cLocationCity)
add_zero <- data.frame(cRace = rep('AS',length(still_miss)),
                       cLocationCity = still_miss,
                       total_count = rep(0,length(still_miss)),
                       Prop = rep(0, length(still_miss)))

#View(AS_row_null)
AS_obs <- brc[brc$cRace == 'AS', ]
AS_obs <- na.omit(AS_obs)

# Verifying positional matching among cities
sum(levels(factor(AS_obs$cLocationCity)) == rownames(AS_row_null))

#View(AS_obs)
AS_row_obs <- AS_obs$Prop

propdif_AS <- AS_row_obs - AS_row_null$AS


ct_to_plot <- city_cts[city_cts$x %in% test_cities2, ]
#View(city_cts)

# Verifying positional matching
ct_to_plot$x == rownames(AS_row_null)

plot(sqrt(ct_to_plot$freq), propdif_AS, main = 'CS: Proportion Difference by Sample Size for AS', xlab = 'Sqrt(City Sample Size)', ylab = 'Proportion Difference (Observed - Census Average)' )

AS_lr <- lm(propdif_AS ~ sqrt(ct_to_plot$freq))
summary(AS_lr)
abline(AS_lr, col = 'red', lwd = 2)

mtext(paste('Adj R-squared=', round(summary(AS_lr)$adj.r.squared,3)), side = 3)

plot(AS_lr, 1)
plot(AS_lr, 4)

```

### CA CS
```{r}
CA_row_null <- as.data.frame(t(null_race_props['CA',]))

# still_miss <- setdiff(rownames(CA_row_null), CA_obs$cLocationCity)
# add_zero <- data.frame(cRace = rep('CA',length(still_miss)),
#                        cLocationCity = still_miss,
#                        total_count = rep(0,length(still_miss)),
#                        Prop = rep(0, length(still_miss)))
# 
# brc <- rbind(brc, add_zero)
#View(CA_row_null)
CA_obs <- brc[brc$cRace == 'CA', ]
CA_obs <- na.omit(CA_obs)

# Verifying positional matching among cities
sum(levels(factor(CA_obs$cLocationCity)) == rownames(CA_row_null))


#View(CA_obs)
CA_row_obs <- CA_obs$Prop

propdif_CA <- CA_row_obs - CA_row_null$CA

ct_to_plot <- city_cts[city_cts$x %in% test_cities2, ]
#View(city_cts)

# Verifying positional matching
ct_to_plot$x == rownames(CA_row_null)

plot(sqrt(ct_to_plot$freq), propdif_CA, main = 'CS: Proportion Difference by Sample Size for CA', xlab = 'Sqrt(City Sample Size)', ylab = 'Proportion Difference (Observed - Census Average)' )

CA_lr <- lm(propdif_CA ~ sqrt(ct_to_plot$freq))
summary(CA_lr)
abline(CA_lr, col = 'red', lwd = 2)

mtext(paste('Adj R-squared=', round(summary(CA_lr)$adj.r.squared,3)), side = 3)

plot(CA_lr, 1)
plot(CA_lr, 4)

```



```{r city counts across total df}
city_cts <- plyr::count(mdd$cLocationCity)
city_cts

# Extracting only races from the null dataframe

null_race_cts <- df_avg[, 1:8]
#View(null_race_cts)
rownames(null_race_cts) <- null_race_cts$City
null_race_cts$City <- NULL

null_race_cts <- t(null_race_cts)

null_race_props <- as.data.frame(apply(null_race_cts, MARGIN = 2, function(x) {x / sum(x)}))

# View(null_race_props)

# count up the subjects in each city by race
brc <- mdd %>%  
  dplyr::group_by(cRace, cLocationCity) %>%  
  dplyr::summarise(total_count=n(), .groups = 'drop')

brc$Prop <- NA
#View(brc)


# turn into proportions
for (i in 1:length(test_cities)) {
  brc$Prop[brc$cLocationCity == test_cities[i]] <- brc$total_count[brc$cLocationCity == test_cities[i]] / sum(brc$total_count[brc$cLocationCity == test_cities[i]])
}
```

### AA MDD
```{r}
AA_row_null <- as.data.frame(t(null_race_props['AA',]))
#View(AA_row_null)
AA_obs <- brc[brc$cRace == 'AA', ]
AA_obs <- na.omit(AA_obs)

# Verifying positional matching among cities
sum(levels(factor(AA_obs$cLocationCity)) == rownames(AA_row_null))


#View(AA_obs)
AA_row_obs <- AA_obs$Prop

propdif_AA <- AA_row_obs - AA_row_null$AA


ct_to_plot <- city_cts[city_cts$x %in% test_cities, ]
#View(city_cts)

# Verifying positional matching
ct_to_plot$x == rownames(AA_row_null)

plot(log(ct_to_plot$freq), propdif_AA, main = 'MDD: Proportion Difference by Sample Size for AA', xlab = 'Log(City Sample Size)', ylab = 'Proportion Difference (Observed - Census Average)' )

AA_lr <- lm(propdif_AA ~ log(ct_to_plot$freq))
summary(AA_lr)
abline(AA_lr, col = 'red', lwd = 2)

mtext(paste('Adj R-squared=', round(summary(AA_lr)$adj.r.squared,3)), side = 3)

plot(AA_lr, 1)
plot(AA_lr, 4)

```

### AS MDD
```{r}
AS_row_null <- as.data.frame(t(null_race_props['AS',]))
#View(AS_row_null)
AS_obs <- brc[brc$cRace == 'AS', ]
AS_obs <- na.omit(AS_obs)

# Verifying positional matching among cities
sum(levels(factor(AS_obs$cLocationCity)) == rownames(AS_row_null))


#View(AS_obs)
AS_row_obs <- AS_obs$Prop

propdif_AS <- AS_row_obs - AS_row_null$AS


ct_to_plot <- city_cts[city_cts$x %in% test_cities, ]
#View(city_cts)

# Verifying positional matching
ct_to_plot$x == rownames(AS_row_null)

plot(log(ct_to_plot$freq), propdif_AS, main = 'MDD: Proportion Difference by Sample Size for AS', xlab = 'Log(City Sample Size)', ylab = 'Proportion Difference (Observed - Census Average)' )

AS_lr <- lm(propdif_AS ~ log(ct_to_plot$freq))
summary(AS_lr)
abline(AS_lr, col = 'red', lwd = 2)

mtext(paste('Adj R-squared=', round(summary(AS_lr)$adj.r.squared,3)), side = 3)

plot(AS_lr, 1)
plot(AS_lr, 4)

```

### CA MDD
```{r}
CA_row_null <- as.data.frame(t(null_race_props['CA',]))
#View(CA_row_null)
CA_obs <- brc[brc$cRace == 'CA', ]
CA_obs <- na.omit(CA_obs)

# Verifying positional matching among cities
sum(levels(factor(CA_obs$cLocationCity)) == rownames(CA_row_null))


#View(CA_obs)
CA_row_obs <- CA_obs$Prop

propdif_CA <- CA_row_obs - CA_row_null$CA

ct_to_plot <- city_cts[city_cts$x %in% test_cities, ]
#View(city_cts)

# Verifying positional matching
ct_to_plot$x == rownames(CA_row_null)

plot(log(ct_to_plot$freq), propdif_CA, main = 'MDD: Proportion Difference by Sample Size for CA', xlab = 'Log(City Sample Size)', ylab = 'Proportion Difference (Observed - Census Average)' )

CA_lr <- lm(propdif_CA ~ log(ct_to_plot$freq))
summary(CA_lr)
abline(CA_lr, col = 'red', lwd = 2)

mtext(paste('Adj R-squared=', round(summary(CA_lr)$adj.r.squared,3)), side = 3)

plot(CA_lr, 1)
plot(CA_lr, 4)