---
title: "Untitled"
author: "Daniel Zoleikhaeian"
date: "2023-03-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r dependencies}
library(plyr)
library(dplyr)
library(readxl)

setwd('C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge')
total_df <- read.csv('C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\bgc_merge_cDiag123_new.csv')
total_df$cEnrollmentYear <- substr(us_data$cEnrollmentDateYear, 1,4)

cogs2 <- total_df[total_df$cStudy == 'COGS2', ]
gpc <- total_df[total_df$cStudy == 'GPC', ]
cogs2_gpc <- rbind(cogs2, gpc)
```

# Table 1: Census Data, GPC, COGS2, and COGS2 + GPC

## Average Census Data
```{r Average Dataset}
df_check_2010 <- read_xlsx('C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\reproj_30_2010.xlsx')
df_check_2020 <- read_xlsx('C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\reproj_30_2020.xlsx')

dem_2010 <- df_check_2010[,-1]
dem_2020 <- df_check_2020[,-1]

#using only 2010 data
df_avg <- df_check_2010



# Reformatting cities
for (i in 1:length(df_avg$City)) {
  df_avg$City[i] <- substr(df_avg$City[i], 1, unlist(gregexpr(',', df_avg$City[i]))[1] - 1)
}

#View(df_avg)


# Standardizing race acronyms
colnames(df_avg)[colnames(df_avg) == 'NA'] <- 'AE'
colnames(df_avg)[colnames(df_avg) == 'NH'] <- 'Not Hispanic'
colnames(df_avg)[colnames(df_avg) == 'PI'] <- 'NH'

# Counts and proportions, formatted for entry into table
race_counts <- df_avg[,1:8]
race_counts <- race_counts[, c('City', 'AA', 'AS', 'AE', 'CA', 'NH', 'MR', 'OT')]

city_tots <- rowSums(race_counts[,-1])
names(city_tots) <- race_counts$City
city_tots

race_counts_only <- race_counts[, -1]
# race_props_2 <- apply(race_counts[-1], MARGIN = 1, function(x) {x / sum(x)})
# colnames(race_props_2) <- df_avg$City
# View(t(race_props_2))

for (i in 1:nrow(race_counts_only)) {
  race_counts_only[i, ] <- race_counts_only[i,] / city_tots[i]
}

# View(race_counts_only)

df_city <- data.frame(City = race_counts$City)

race_props <- cbind(df_city, race_counts_only)
View(race_props)

city_totals <- apply(race_counts[,-1], MARGIN = 1, sum)

race_counts_w_tot <- cbind(race_counts, city_totals)
colnames(race_counts_w_tot)[ncol(race_counts_w_tot)] <- 'Total'

# write.csv(race_counts, file = 'census_avg_race_n.csv')
# write.csv(race_props, file = 'census_avg_race_props.csv')
# write.csv(race_counts_w_tot, file = 'census_avg_race_nw_tot.csv')
```

## 2010 and 2020 census data
```{r raw counts 2010 and 2020}
df_2010 <- read_xlsx('reproj_30_2010.xlsx')[, 1:8]
df_2010 <- df_2010[, c('City', 'AA', 'AS', 'NA', 'CA', 'PI', 'MR', 'OT')]
write.csv(df_2010, file = 'n_df2010_reord.csv')

df_2020 <- read_xlsx('reproj_30_2020.xlsx')[, 1:8]
df_2020 <- df_2020[, c('City', 'AA', 'AS', 'NA', 'CA', 'PI', 'MR', 'OT')]
write.csv(df_2020, file = 'n_df2020_reord.csv')
```

```{r proportions 2010 and 2020}
df_2010_props <- apply(df_2010[-1], MARGIN = 1, function(x) {x / sum(x)})
df_2010_props <- as.data.frame(t(df_2010_props))
rownames(df_2010_props) <- df_avg$City
write.csv(df_2010_props, file = '2010_census_props.csv')

df_2020_props <- apply(df_2020[-1], MARGIN = 1, function(x) {x / sum(x)})
df_2020_props <- as.data.frame(t(df_2020_props))
rownames(df_2020_props) <- df_avg$City
write.csv(df_2020_props, file = '2020_census_props.csv')
```

## Proportions pipeline
```{r}
prop_fn <- function(study, total_ds, city_nm, diag4) {
  ds_subset <- total_ds[total_ds$cStudy %in% study & total_ds$cLocationCity == city_nm & total_ds$cDiagnosis4 %in% diag4, ]
  
  ## Converting all races from OT or UNK into OT/UNK
  ds_subset$cRace[ds_subset$cRace == 'OT' | ds_subset$cRace == 'UNK'] <- 'OT/UNK'
  
  race_cts <- plyr::count(ds_subset, 'cRace')
  
  race_cts$prop <- race_cts$freq / sum(race_cts$freq)
  race_cts <- t(race_cts)
  
  # for some reason, r converts the dataframe into a vector if you delete a row with only one element
  if (length(race_cts[1, ]) == 1) {
    race_name <- race_cts['cRace', 1]
    race_cts <- as.data.frame(race_cts[-1])
    colnames(race_cts) <- race_name
  } else {
    colnames(race_cts) <- race_cts['cRace', ]
    race_cts <- race_cts[-1, ]
  }
  
  total_races <- c('AA', 'AS', 'AE', 'CA', 'NH', 'MR', 'OT/UNK')
  
  not_in <- which(!(total_races %in% colnames(race_cts)))
  
  if (length(not_in) > 0) {
    for (i in 1:length(not_in)) {
      zero_df <- data.frame(c(0, 0))
      race_cts <- cbind(race_cts, zero_df)
    }
  
    colnames(race_cts)[(ncol(race_cts) - length(not_in) + 1):ncol(race_cts)] <- total_races[not_in]
  }
  
  #path <- 'C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\012623\\Data Table City Proportions\\City Proportions\\'
  path <- 'C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\Table 1 - Data Table City Proportions\\City Proportions - All and BSNIP\\'
  
  if (length(diag4) > 1) {
    cond_name <- 'All'
  }
  else {
    cond_name <- diag4
  }
  
  
  full_path <- paste(path, paste(city_nm, cond_name, 'race', paste(study, collapse = '-'), '.csv', sep = '_'))
  
  race_cts <- race_cts[, total_races]
  
  if (sum(as.numeric(race_cts[2, ])) == 0) {
    print(paste(city_nm, 'does not have any individuals for', paste(study, collapse = '-'), 'and Diagnosis', cond_name))
    return ()
  }
  write.csv(race_cts, file = full_path, row.names = F)
  return (race_cts)
}
```

## Looping through individual conditions, the study combinations, and cities
```{r }

cities_to_examine <- levels(factor(total_df$cLocationCity)) %in% df_avg$City
cities <- levels(factor(total_df$cLocationCity))[cities_to_examine]
conditions <- list(levels(factor(total_df$cDiagnosis4)), 'CS', 'SZSAFD', 'BAD12', 'MDD')

for (i in 1:length(cities)) {
  for(j in 1:length(conditions)) {
    gpc_df <- prop_fn('GPC', total_df, cities[i], diag4 = conditions[[j]])
    cogs2_df <- prop_fn('COGS2', total_df, cities[i], diag4 = conditions[[j]])
    cogs2_gpc_df <- prop_fn(c('GPC', 'COGS2'), total_df, cities[i], diag4 = conditions[[j]])
  }
}
```

## Adding BSNIP1 and BSNIP2
```{r}
for (i in 1:length(cities)) {
  for(j in 1:length(conditions)) {
    all_df <- prop_fn(unique(total_df$cStudy), total_df, cities[i], diag4 = conditions[[j]])
    bsnip_df <- prop_fn(c('BSNIP1', 'BSNIP2'), total_df, cities[i], diag4 = conditions[[j]])
  }
}


```

## Making university proportions

```{r}
uni_props_fn <- function(df_in, studies_, diag4) {
  
  res_df_list() <- list()
  
  for (k in 1:length(studies_)) {
    
    ds_subset <- df_in[df_in$cStudy %in% studies_ & df_in$cDiagnosis4 %in% diag4, ]
  cities_ <- unique(ds_subset$cLocationCity)

    ds_subset$cRace[ds_subset$cRace[ds_subset$cRace == 'OT' | ds_subset$cRace == 'UNK'] <- 'OT/UNK']
    
    
    for (i in 1:length(cities_)) {
      
      uni_counts <- plyr::count(ds_subset[ds_subset$cLocationCity == cities_[i], ], 'cLocationInstitution')
      if (nrow(uni_counts) == 0) {
        next
      }
      
      uni_df1 <- ds_subset[ds_subset$cLocationInstitution == uni_counts$cLocationInstitution[1], ]
      uni_df1_race_counts <- plyr::count(uni_df1, 'cRace')
      uni_df1_race_props <- uni_df1_race_counts$freq / sum(uni_df1_race_counts$freq)
      names(uni_df1_race_props) <- uni_df1_race_counts$cRace
      
      
      uni_df_race_props_df <- as.data.frame(t(uni_df1_race_props), row.names = uni_counts$cLocationInstitution[1])
      
      # adding the missing races
      
      not_in_races_bool <- !(total_races %in% colnames(uni_df_race_props_df))
      not_in_races <- total_races[not_in_races_bool]
      
      if (length(not_in_races) > 0) {
        for (j in 1:length(not_in_races)) {
        
        df_to_bind <- data.frame(race_hold = 0, row.names = uni_counts$cLocationInstitution[1])
        
        uni_df_race_props_df <- cbind(uni_df_race_props_df, df_to_bind)
      }
      
        colnames(uni_df_race_props_df)[(ncol(uni_df_race_props_df) - length(not_in_races) + 1):ncol(uni_df_race_props_df)] <- not_in_races
      }
      
      uni_df_race_props_df <- uni_df_race_props_df[, total_races]
    
      
      if (nrow(uni_counts) > 1) {
        uni_df2 <- ds_subset[ds_subset$cLocationInstitution == uni_counts$cLocationInstitution[2], ]
        uni_df2_race_counts <- plyr::count(uni_df2, 'cRace')
        uni_df2_race_props <- uni_df2_race_counts$freq / sum(uni_df2_race_counts$freq)
        names(uni_df2_race_props) <- uni_df2_race_counts$cRace
        
        uni_df_race_props_df2 <- as.data.frame(t(uni_df2_race_props), row.names = uni_counts$cLocationInstitution[2])
        
        # adding the missing races
        
        not_in_races_bool <- !(total_races %in% colnames(uni_df_race_props_df2))
        not_in_races <- total_races[not_in_races_bool]
        if (length(not_in_races) > 0) {
        
            for (j in 1:length(not_in_races)) {
              
              df_to_bind <- data.frame(race_hold = 0, row.names = uni_counts$cLocationInstitution[2])
              
              uni_df_race_props_df2 <- cbind(uni_df_race_props_df2, df_to_bind)
            }
          
            colnames(uni_df_race_props_df2)[(ncol(uni_df_race_props_df2) - length(not_in_races) + 1):ncol(uni_df_race_props_df2)] <- not_in_races
          }
          
        
          uni_df_race_props_df2 <- uni_df_race_props_df2[, total_races]
            
          uni_df_race_props_df <- rbind(uni_df_race_props_df, uni_df_race_props_df2)
          rownames(uni_df_race_props_df)[2] <- uni_counts$cLocationInstitution[2]
      }
    }
    
    res_df_list[[k]] <- res_df_out
  }
  
  path_prefix <- 'C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\University Proportions\\'
      
      if (length(diag4) > 1) {
        diag_alias <- 'All' 
      } else {
        diag_alias <- diag4
      }
      
      path_suffix <- paste(diag_alias, '_', paste(studies_, collapse = '+'), '_uni_props', cities_[i], '.csv', sep = '')
      path_full <- paste(path_prefix, path_suffix, sep = '')
      
      write.csv(uni_df_race_props_df, file = path_full)
  
  
  

  
}

```

```{r}
uni_props_fn_og <- function(df_in, studies_, diag4) {
  
  ds_subset <- df_in[df_in$cStudy %in% studies_ & df_in$cDiagnosis4 %in% diag4, ]
  cities_ <- unique(ds_subset$cLocationCity)

  ds_subset$cRace[ds_subset$cRace[ds_subset$cRace == 'OT' | ds_subset$cRace == 'UNK'] <- 'OT/UNK']
  
  
  for (i in 1:length(cities_)) {
    
    uni_counts <- plyr::count(ds_subset[ds_subset$cLocationCity == cities_[i], ], 'cLocationInstitution')
    
    if (nrow(uni_counts) == 0) {
      next
    }
    
    uni_counts <- uni_counts[order(-uni_counts$freq), ]
    
    uni_df1 <- ds_subset[ds_subset$cLocationInstitution == uni_counts$cLocationInstitution[1], ]
    uni_df1_race_counts <- plyr::count(uni_df1, 'cRace')
    uni_df1_race_props <- uni_df1_race_counts$freq / sum(uni_df1_race_counts$freq)
    names(uni_df1_race_props) <- uni_df1_race_counts$cRace
    
    
    uni_df_race_props_df <- as.data.frame(t(uni_df1_race_props), row.names = uni_counts$cLocationInstitution[1])
    
    # adding the missing races
    
    not_in_races_bool <- !(total_races %in% colnames(uni_df_race_props_df))
    not_in_races <- total_races[not_in_races_bool]
    
    if (length(not_in_races) > 0) {
      for (j in 1:length(not_in_races)) {
      
      df_to_bind <- data.frame(race_hold = 0, row.names = uni_counts$cLocationInstitution[1])
      
      uni_df_race_props_df <- cbind(uni_df_race_props_df, df_to_bind)
    }
    
      colnames(uni_df_race_props_df)[(ncol(uni_df_race_props_df) - length(not_in_races) + 1):ncol(uni_df_race_props_df)] <- not_in_races
    }
    
    uni_df_race_props_df <- uni_df_race_props_df[, total_races]
  
    
    if (nrow(uni_counts) > 1) {
      uni_df2 <- ds_subset[ds_subset$cLocationInstitution == uni_counts$cLocationInstitution[2], ]
      uni_df2_race_counts <- plyr::count(uni_df2, 'cRace')
      uni_df2_race_props <- uni_df2_race_counts$freq / sum(uni_df2_race_counts$freq)
      names(uni_df2_race_props) <- uni_df2_race_counts$cRace
      
      uni_df_race_props_df2 <- as.data.frame(t(uni_df2_race_props), row.names = uni_counts$cLocationInstitution[2])
      
      # adding the missing races
      
      not_in_races_bool <- !(total_races %in% colnames(uni_df_race_props_df2))
      not_in_races <- total_races[not_in_races_bool]
      if (length(not_in_races) > 0) {
      
          for (j in 1:length(not_in_races)) {
            
            df_to_bind <- data.frame(race_hold = 0, row.names = uni_counts$cLocationInstitution[2])
            
            uni_df_race_props_df2 <- cbind(uni_df_race_props_df2, df_to_bind)
          }
        
          colnames(uni_df_race_props_df2)[(ncol(uni_df_race_props_df2) - length(not_in_races) + 1):ncol(uni_df_race_props_df2)] <- not_in_races
        }
        
      
        uni_df_race_props_df2 <- uni_df_race_props_df2[, total_races]
          
        uni_df_race_props_df <- rbind(uni_df_race_props_df, uni_df_race_props_df2)
        rownames(uni_df_race_props_df)[2] <- uni_counts$cLocationInstitution[2]
    }
      
    path_prefix <- 'C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\University Proportions\\'
    
    if (length(diag4) > 1) {
      diag_alias <- 'All' 
    } else {
      diag_alias <- diag4
    }
    
    path_suffix <- paste(diag_alias, '_', cities_[i], paste(studies_, collapse = '+'), '_uni_props',  '.csv', sep = '')
    path_full <- paste(path_prefix, path_suffix, sep = '')
    
    write.csv(uni_df_race_props_df, file = path_full)
    
  }
  
}
```

## Uni props function call
```{r}

# variables to loop through
study_combos <- list('GPC', 'COGS2', c('GPC', 'COGS2'))
cond_list <- list(unique(total_df$cDiagnosis4), 'CS', 'SZSAFD', 'BAD12', 'MDD')

for (i in 1:length(study_combos)) {
  for (j in 1:length(cond_list)) {
    uni_props_fn_og(total_df, study_combos[[i]], cond_list[[j]])
  }
}
 


```

## Uni props tester pipeline
```{r}
total_races <- c('AA', 'AS', 'AE', 'CA', 'NH', 'MR', 'OT/UNK')

cities_cogs <- unique(cogs2$cLocationCity)

cogs2$cRace[cogs2$cRace[cogs2$cRace == 'OT' | cogs2$cRace == 'UNK'] <- 'OT/UNK']


for (i in 1:length(cities_cogs)) {
  
  uni_counts <- plyr::count(cogs2[cogs2$cLocationCity == cities_cogs[i], ], 'cLocationInstitution')
  #View(uni_counts)
  
  uni_df1 <- cogs2[cogs2$cLocationInstitution == uni_counts$cLocationInstitution[1], ]
  uni_df1_race_counts <- plyr::count(uni_df1, 'cRace')
  uni_df1_race_props <- uni_df1_race_counts$freq / sum(uni_df1_race_counts$freq)
  names(uni_df1_race_props) <- uni_df1_race_counts$cRace
  
  
  uni_df_race_props_df <- as.data.frame(t(uni_df1_race_props), row.names = uni_counts$cLocationInstitution[1])
  
  # adding the missing races
  
  not_in_races_bool <- !(total_races %in% colnames(uni_df_race_props_df))
  not_in_races <- total_races[not_in_races_bool]
  
  if (length(not_in_races) > 0) {
    for (j in 1:length(not_in_races)) {
    
    df_to_bind <- data.frame(race_hold = 0, row.names = uni_counts$cLocationInstitution[1])
    
    uni_df_race_props_df <- cbind(uni_df_race_props_df, df_to_bind)
  }
  
    colnames(uni_df_race_props_df)[(ncol(uni_df_race_props_df) - length(not_in_races) + 1):ncol(uni_df_race_props_df)] <- not_in_races
  }
  
  
  
  uni_df_race_props_df <- uni_df_race_props_df[, total_races]

  
  #View(uni_df_race_props_df)
  
  
  if (nrow(uni_counts) > 1) {
    uni_df2 <- cogs2[cogs2$cLocationInstitution == uni_counts$cLocationInstitution[2], ]
    uni_df2_race_counts <- plyr::count(uni_df2, 'cRace')
    uni_df2_race_props <- uni_df2_race_counts$freq / sum(uni_df2_race_counts$freq)
    names(uni_df2_race_props) <- uni_df2_race_counts$cRace
    
    uni_df_race_props_df2 <- as.data.frame(t(uni_df2_race_props), row.names = uni_counts$cLocationInstitution[2])
    
    # adding the missing races
    if (length(not_in_races) > 0) {
      not_in_races_bool <- !(total_races %in% colnames(uni_df_race_props_df2))
  not_in_races <- total_races[not_in_races_bool]
  
  for (j in 1:length(not_in_races)) {
    
    df_to_bind <- data.frame(race_hold = 0, row.names = uni_counts$cLocationInstitution[2])
    
    uni_df_race_props_df2 <- cbind(uni_df_race_props_df2, df_to_bind)
  }
  
    colnames(uni_df_race_props_df2)[(ncol(uni_df_race_props_df2) - length(not_in_races) + 1):ncol(uni_df_race_props_df2)] <- not_in_races
    }
    
  
  uni_df_race_props_df2 <- uni_df_race_props_df2[, total_races]
    
    uni_df_race_props_df <- rbind(uni_df_race_props_df, uni_df_race_props_df2)
    rownames(uni_df_race_props_df)[2] <- uni_counts$cLocationInstitution[2]
  }
  
  path_prefix <- 'C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\University Proportions\\'
  path_full <- paste(path_prefix, 'COGS2_uni_props', cities_cogs[i], '.csv', sep = '')
  
  write.csv(uni_df_race_props_df, file = path_full)
  
}





```

# Table 2: Proportion Differences (Observed - Census Average)

## Proportion Differences (Observed - Census Average)

### Dependency: Finding Proportion Differences
```{r}
race_cities_plot_data2 <- function (null_df, tot_df, disease, study, race_acr) {
  
  obs_df <- tot_df[tot_df$cStudy %in% study & tot_df$cDiagnosis4 %in% disease, ]
  
  obs_cities <- obs_df$cLocationCity[obs_df$cRace == race_acr]
  
  test_cities <- intersect(null_df$City, obs_cities)
  
  null_race_cts <- null_df[, 1:8]

  null_race_cts <- df_avg[, 1:8]
  null_race_props <- null_race_cts

  for (i in 1:nrow(null_race_cts)) {
    city_tot <- sum(null_race_cts[i, 2:8])
    null_race_props[i, 2:8] <- null_race_cts[i,2:8] / city_tot
  }
  null_race_props <- null_race_props[null_race_props$City %in% test_cities,]
  null_race_props <- null_race_props[order(null_race_props$City), ]
  
  brc <- obs_df %>%  
    dplyr::group_by(cRace, cLocationCity) %>%  
    dplyr::summarise(total_count=n(), .groups = 'drop')
  
  brc <- brc[brc$cLocationCity %in% test_cities, ]
  
  brc <- brc[order(brc$cLocationCity), ]
  brc$Prop <- NA
  
  # Turning counts into proportions
  for (i in 1:length(test_cities)) {
    brc$Prop[brc$cLocationCity == test_cities[i]] <- brc$total_count[brc$cLocationCity == test_cities[i]] / sum(brc$total_count[brc$cLocationCity == test_cities[i]])
  }
  
  null_row <- as.data.frame(null_race_props[,c('City', race_acr)])
  obs_races <- brc[brc$cRace == race_acr, ]
  obs_prop <- obs_races$Prop
  
  ## Begin output variables to send to lm() and plot()
  propdif <- obs_prop - null_row[,2] ## response
  names(propdif) <- test_cities
  
  city_cts <- plyr::count(obs_df$cLocationCity)
  ct_to_plot <- city_cts$freq[city_cts$x %in% test_cities] ## predictor 1
  
  null_race_cts <- null_race_cts[null_race_cts$City %in% test_cities, ]
  race_of_interest_null <- null_race_cts[, colnames(null_race_cts) == race_acr] ## predictor 2
  
  null_race_sums <- as.data.frame(apply(null_race_cts[,-1], 1, sum)) ## predictor 3
  colnames(null_race_sums) <- 'Total City Population'
  
  output <- list(propdif, ct_to_plot, race_of_interest_null, null_race_sums[,1])
  names(output) <- c('Prop dif', 'city sample', 'census average race count', 'census average pop total')
  
  return (output)
  
}
```


### Proportion Deltas

### Pre-Processing for Proportion Deltas Pipeline
```{r}
total_df_copy <- total_df[total_df$cEnrollmentYear %in% 2007:2013, ]
total_df_copy$cRace[total_df_copy$cRace == 'OT/UNK' | total_df_copy$cRace == 'UNK'] <- 'OT'
race_acrs <- c('AA', 'AS', 'AE', 'CA', 'NH', 'MR', 'OT')
city_vec <- df_avg$City
cond_list <- list(unique(total_df$cDiagnosis4), 'CS', 'SZSAFD', 'BAD12', 'MDD')
df_tot_props <- data.frame(matrix(, nrow=length(city_vec), ncol=0))
rownames(df_tot_props) <- city_vec

```

### Proportion Deltas Pipeline
```{r}
for (i in 1:length(cond_list)) {
  for (j in 1:length(race_acrs)) {
    deltas <- race_cities_plot_data2(df_avg, total_df_copy, disease = cond_list[[i]], study = c('GPC', 'COGS2'), race_acr = race_acrs[j])
    prop_difs <- as.data.frame(deltas$`Prop dif`)
    
    not_in <- which(!(city_vec %in% rownames(prop_difs)))
  city_vec[not_in]
  
    zero_vec <- rep(0, length(not_in))
    names(zero_vec) <- city_vec[not_in]
    
    zero_vec_df <- (data.frame(zero_vec))
    colnames(zero_vec_df) <- 'hold'
    colnames(prop_difs) <- 'hold'
    
    prop_difs <- rbind(prop_difs, zero_vec_df)
    
    #View(prop_difs)
    
    prop_difs <- prop_difs[city_vec, ]
    names(prop_difs) <- city_vec
    prop_difs <- data.frame(prop_difs)
    colnames(prop_difs) <- race_acrs[j]
    
    df_tot_props <- cbind(df_tot_props, prop_difs)
  }
}

write.csv(df_tot_props, file = 'C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\Table 2 - Proportion Deltas\\prop_delts_cogsgpc_041023.csv', row.names = F)
```



#### Tester case 2
```{r}

for (i in 1:length(race_acrs)) {
  deltas <- race_cities_plot_data2(df_avg, total_df_copy, disease = unique(total_df$cDiagnosis4), study = c('GPC', 'COGS2'), race_acr = race_acrs[i])
  prop_difs <- as.data.frame(deltas$`Prop dif`)
  
  not_in <- which(!(city_vec %in% rownames(prop_difs)))
city_vec[not_in]

  zero_vec <- rep(0, length(not_in))
  names(zero_vec) <- city_vec[not_in]
  
  zero_vec_df <- (data.frame(zero_vec))
  colnames(zero_vec_df) <- 'hold'
  colnames(prop_difs) <- 'hold'
  
  prop_difs <- rbind(prop_difs, zero_vec_df)
  
  #View(prop_difs)
  
  prop_difs <- prop_difs[city_vec, ]
  names(prop_difs) <- city_vec
  prop_difs <- data.frame(prop_difs)
  colnames(prop_difs) <- race_acrs[i]
  
  df_tot_props <- cbind(df_tot_props, prop_difs)

}
View(df_tot_props)

write.csv(df_tot_props, file = 'C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\Table 2 - Proportion Deltas\\prop_delts_cogsgpc_041023.csv', row.names = F)
```

#### Proportion Deltas Pipeline with All Studies
```{r}

city_vec <- df_avg$City
cond_list <- list(unique(total_df$cDiagnosis4), 'CS', 'SZSAFD', 'BAD12', 'MDD')
df_tot_props_all <- data.frame(matrix(, nrow=length(city_vec), ncol=0))
rownames(df_tot_props_all) <- city_vec

for (i in 1:length(cond_list)) {
  for (j in 1:length(race_acrs)) {
    deltas <- race_cities_plot_data2(df_avg, total_df_copy, disease = cond_list[[i]], study = c('GPC', 'COGS2', 'BSNIP1', 'BSNIP2'), race_acr = race_acrs[j])
    prop_difs <- as.data.frame(deltas$`Prop dif`)
    
    not_in <- which(!(city_vec %in% rownames(prop_difs)))
  city_vec[not_in]
  
    zero_vec <- rep(0, length(not_in))
    names(zero_vec) <- city_vec[not_in]
    
    zero_vec_df <- (data.frame(zero_vec))
    colnames(zero_vec_df) <- 'hold'
    colnames(prop_difs) <- 'hold'
    
    prop_difs <- rbind(prop_difs, zero_vec_df)
    
    #View(prop_difs)
    
    prop_difs <- prop_difs[city_vec, ]
    names(prop_difs) <- city_vec
    prop_difs <- data.frame(prop_difs)
    colnames(prop_difs) <- race_acrs[j]
    
    df_tot_props_all <- cbind(df_tot_props_all, prop_difs)
  }

}

write.csv(df_tot_props_all, file = 'C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\Table 2 - Proportion Deltas\\prop_delts_all_040523.csv', row.names = F)



```



# Table 3: Regression Results for COGS2 + GPC
```{r}
studies <- c('COGS2', 'GPC')
out_list <- list()
lm_results_df <- data.frame(matrix(NA, nrow = 12, ncol = 0))

for (i in 1:length(cond_list)) {
  for (j in 1:length(race_acrs)) {
    
    in_list <- race_cities_plot_data2(df_avg, total_df_copy, cond_list[[i]], studies, race_acrs[j])
    
    out_list[[length(out_list) + 1]] <- in_list
    if (length(cond_list[[i]]) == 1) {
      names(out_list)[length(out_list)] <- paste(cond_list[[i]], race_acrs[j])
    } else {
      names(out_list)[length(out_list)] <- paste('All', race_acrs[j])
    }
    
    out_vec <- c()
    for (k in 2:4) {
      if (length(in_list[[1]]) < 1) {
        #out_vec <- rep(NA, 21)
        out_vec <- rep(NA, 12)
        next
      }
      
      lm_res <- lm(in_list[[1]] ~ log(unlist(in_list[[k]])) - 1)
      summary_lm_res <- summary(lm_res)
      int_coeff_std_pval <- summary_lm_res$coefficients[1, c(1, 2, 4)]

      
      # if (length(in_list[[1]]) == 1) {
      #   pred_coeff_std_pval <- rep(NA, 3)
      # } else {
      #   pred_coeff_std_pval <- summary_lm_res$coefficients[2, c(1, 2, 4)]
      # }
      
      #r_squared <- summary_lm_res$adj.r.squared
      r_squared <- summary_lm_res$r.squared
      #out_vec <- c(out_vec, int_coeff_std_pval, pred_coeff_std_pval, r_squared)
      out_vec <- c(out_vec, int_coeff_std_pval, r_squared)
    }
    
    lm_results_df <- cbind(lm_results_df, out_vec)
  }
}

colnames(lm_results_df) <- rep(race_acrs, 5)

View(lm_results_df)

write.csv(lm_results_df, file = 'C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\Table 3 - Linear Regression Results\\cogsgpc07013_no_int_r2_041723.csv', row.names = F)

```

# Table 3: Regression Results for COGS2 + GPC + BSNIP12
```{r}
studies <- c('COGS2', 'GPC', 'BSNIP1', 'BSNIP2')
out_list <- list()
#lm_results_df <- data.frame(matrix(NA, nrow=21, ncol=0))
lm_results_df <- data.frame(matrix(NA, nrow = 12, ncol = 0))

for (i in 1:length(cond_list)) {
  for (j in 1:length(race_acrs)) {
    
    in_list <- race_cities_plot_data2(df_avg, total_df_copy, cond_list[[i]], studies, race_acrs[j])
    
    out_list[[length(out_list) + 1]] <- in_list
    if (length(cond_list[[i]]) == 1) {
      names(out_list)[length(out_list)] <- paste(cond_list[[i]], race_acrs[j])
    } else {
      names(out_list)[length(out_list)] <- paste('All', race_acrs[j])
    }
    
    out_vec <- c()
    for (k in 2:4) {
      if (length(in_list[[1]]) < 1) {
        #out_vec <- rep(NA, 21)
        out_vec <- rep(NA, 12)
        next
      }
      
      lm_res <- lm(in_list[[1]] ~ log(unlist(in_list[[k]])) - 1)
      summary_lm_res <- summary(lm_res)
      int_coeff_std_pval <- summary_lm_res$coefficients[1, c(1, 2, 4)]

      
      # if (length(in_list[[1]]) == 1) {
      #   pred_coeff_std_pval <- rep(NA, 3)
      # } else {
      #   pred_coeff_std_pval <- summary_lm_res$coefficients[2, c(1, 2, 4)]
      # }
      
      #r_squared <- summary_lm_res$adj.r.squared
      r_squared <- summary_lm_res$r.squared
      #out_vec <- c(out_vec, int_coeff_std_pval, pred_coeff_std_pval, r_squared)
      out_vec <- c(out_vec, int_coeff_std_pval, r_squared)
    }
    
    lm_results_df <- cbind(lm_results_df, out_vec)
  }
}

colnames(lm_results_df) <- rep(race_acrs, 5)

View(lm_results_df)

write.csv(lm_results_df, file = 'C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\Table 3 - Linear Regression Results\\all07013_no_int_r2_041723.csv', row.names = F)

```
