---
title: "27. 2010-2020 Race Proportion Differences by City"
author: "Daniel Zoleikhaeian"
date: "2023-04-19"
output: pdf_document
---

# knit set-up
```{r}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy = TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=60),tidy=TRUE)
```

# Dependencies
```{r}
library(plyr)
library(dplyr)
library(ggplot2)
library(readxl)
```

# Datasets
```{r}
dat_2010 <- read_xlsx('C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\reproj_30_2010.xlsx')
dat_2020 <- read_xlsx('C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\reproj_30_2020.xlsx')
View(dat_2010)
View(dat_2020)

dat_2010_races <- dat_2010[,2:8]
dat_2020_races <- dat_2020[,2:8]

rownames(dat_2010_races) <- dat_2010$City
rownames(dat_2020_races) <- dat_2020$City
View(dat_2010_races)
```

# Making proportions
```{r}

# Making proportions
prop_2010 <- as.data.frame(t(apply(dat_2010_races, MARGIN = 1, function(x) {x / sum(x)}, simplify = T)))

# Making a city column
prop_2010$City <- rownames(prop_2010)
prop_2010 <- prop_2010[, c(ncol(prop_2010), 1:(ncol(prop_2010) - 1))]
rownames(prop_2010) <- NULL

# reformatting the City vector
prop_2010$City <- lapply(prop_2010$City, function (x)  {substr(x, 1, unlist(gregexpr(',', x))[1] - 1)}  )

# Adding a Year column
prop_2010$Year <- rep(2010, nrow(prop_2010))

prop_2020 <- as.data.frame(t(apply(dat_2020_races, MARGIN = 1, function(x) {x / sum(x)}, simplify = T)))

prop_2020$City <- rownames(prop_2020)
prop_2020 <- prop_2020[, c(ncol(prop_2020), 1:(ncol(prop_2020) - 1))]
rownames(prop_2020) <- NULL

prop_2020$City <- lapply(prop_2020$City, function (x)  {substr(x, 1, unlist(gregexpr(',', x))[1] - 1)}  )

prop_2020$Year <- rep(2020, nrow(prop_2020))

# stacking the dataframes on top of each other
prop_2010_2020 <- rbind(prop_2010, prop_2020)

# checking if the rows are equal
nrow(prop_2010_2020) == (nrow(prop_2010) + nrow(prop_2020))

# standardizing acronyms
colnames(prop_2010_2020)[colnames(prop_2010_2020) == 'NA'] <- 'AE'
colnames(prop_2010_2020)[colnames(prop_2010_2020) == 'PI'] <- 'NH'
```

# Visualization 1: Plots of the proportions in 2010 and 2020
```{r}
# Plots for each race
# Lines for each city

prop_2010_2020$City <- factor(as.character(prop_2010_2020$City))

race_acrs <- c('AA', 'AS', 'CA', 'AE', 'NH', 'MR', 'OT')

plot_race_cities <- function(race, data_in) {
  ggplot(data_in, aes_string(x="Year",
                      y=race,
                      col="City")) +
  geom_point() + 
  ylab('Proportion') +
  geom_smooth(method=lm, se = FALSE) + 
  ggtitle(paste('Race Proportions for', race))
}

plot_collection <- lapply(race_acrs, plot_race_cities, data_in = prop_2010_2020)

# plot_collection
```

Conclusions:
* Race proportions not stable over time
* Shifts generally less than 0.05 for most cities

# Hypothesis Testing: Proportion Difference between 2010 and 2020

## Calculating proportion differences between 2010 and 2020
```{r}

# Calculated as proportion in 2020 - proportion in 2010

prop_difs <- prop_2020[,2:8] - prop_2010[,2:8]
prop_difs$City <- prop_2020$City

# standardizing acronyms
colnames(prop_difs)[colnames(prop_difs) == 'NA'] <- 'AE'
colnames(prop_difs)[colnames(prop_difs) == 'PI'] <- 'NH'

# re-ordering columns
prop_difs <- prop_difs[, c('City', race_acrs)]

# making sub-dataframes for AA, AS, and CA
AA_df <- prop_difs[, c('City', 'AA')]
AA_df$p_val <- rep(-1, nrow(AA_df))

AS_df <- prop_difs[, c('City', 'AS')]
AS_df$p_val <- rep(-1, nrow(AS_df))

CA_df <- prop_difs[, c('City', 'CA')]
CA_df$p_val <- rep(-1, nrow(CA_df))
```

## Visualization 2: Proportion Differences of Races across Cities
```{r}

bar_plot_race_cities <- function(race_acr, data_in) {
  ggplot(data = data_in, aes_string(x="City",y=race_acr, color="City")) +
    geom_bar(stat='identity', fill = 'white') +
    ylab('Difference (2020 - 2010)') +
    ggtitle(paste(race_acr, ':Proportion Difference between 2010 and 2020'))+ 
    xlab('City') + 
    coord_flip() +
    theme(legend.position="none")
}

prop_difs$City <- as.factor(unlist(prop_difs$City))

plot_collection2 <- lapply(race_acrs, bar_plot_race_cities, data_in = prop_difs)

# plot_collection2

## Running plot collections side by side
for (i in 1:7) {
  print(plot_collection[[i]])
  print(plot_collection2[[i]])
}

```

## Testing if proportion differences are significant for each city
```{r}
dat_2010_races$City_n <- rowSums(dat_2010_races)
dat_2020_races$City_n <- rowSums(dat_2020_races)

dat_2010_races <- as.data.frame(dat_2010_races)
dat_2020_races <- as.data.frame(dat_2020_races)

# Probabilities deviate far from 0.5, so do not use Z-test
# Use Fisher exact test instead
## Benefits: Works with small counts
## Drawbacks: Conservative; hard to reject null hypothesis

## EXAMPLE: AA in Athens between 2010 and 2020
df <- data.frame(AA_yes = c(dat_2010_races[1, 'AA'], dat_2020_races[1, 'AA']),
                 AA_no = c(dat_2010_races$City_n[1] - dat_2010_races[1, 'AA'], dat_2020_races$City_n[1] - dat_2020_races[1, 'AA']), 
                 row.names = c('2010', '2020'))
View(df)
res <- fisher.test(df)
res$p.value

# Looping through all the columns of AA, AS, and CA

add_fish_p_val <- function(race_acr) {
  
  p_vals <- rep(0, nrow(dat_2010_races))
  
  for (i in 1:nrow(dat_2010_races)) {
    df <- data.frame(Race_yes = c(dat_2010_races[i, race_acr], dat_2020_races[i, race_acr]),
                 Race_no = c(dat_2010_races$City_n[i] - dat_2010_races[i, race_acr], dat_2020_races$City_n[i] - dat_2020_races[i, race_acr]), 
                 row.names = c('2010', '2020'))
   res <- fisher.test(df)
   p_vals[i] <- res$p.value
  }
  
  
  return (p_vals)
    
}

AA_df$p_val <- add_fish_p_val('AA')
AS_df$p_val <- add_fish_p_val('AS')
CA_df$p_val <- add_fish_p_val('CA')

```

## Dataframe of proportion differences and p-values
```{r}
## African-Americans
AA_df
```

```{r}
## Asian-Americans
AS_df
```


```{r}
## Caucasian-Americans
CA_df
```

Conclusions:

* Magnitude of differences is small for AS and AA
  * All differences are significant at the 0.05 level for AS
  * Some are not significant for AA
  * High power to detect small difference due to large sample size
  
* Magnitude of difference for CA much higher than for AS and AA
  * General decrease in proportion of CA between 2010 and 2020
  * All differences are significant

* Suggestion:
  * Designate a magnitude and p-value cutoff to determine which differences are consequential for the census "average"

# Final Thoughts:
* Cannot use "average" 2010 and 2020 for the null hypothesis
  * 2010 and 2020 are different, but magnitude may not be consequential

