---
title: "Proportions over Time"
author: "Daniel Zoleikhaeian"
date: "2023-04-24"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries
```{r}
library(plyr)
library(dplyr)
library(ggplot2)
library(readxl)
```

## Loading Dataset
```{r}
us_data <- read.csv('C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\bgc_merge_cDiag123_new.csv')
head(us_data)
us_data$cEnrollmentYear <- substr(us_data$cEnrollmentDateYear, 1,4)

# Combining OT/UNK and UNK into OT/UNK
us_data$cRace[us_data$cRace %in% c('UNK', 'OT/UNK')] <- 'OT'

# Setting working directory
setwd("C:/Users/danie/Documents/Joshi Lab Materials/3 Studies Dataset/Dataset Merge/Rmd Files/Enrollment Plots")
```

## Pipeline for calculating proportions over time - not cumulative
```{r}
props_over_time <- function(studies) {
  
  # make dataframe to store results
  prop_res <- data.frame(Race = c('AA', 'AE', 'AS', 'CA', 'MR', 'NH', 'OT'),
                         y2010 = rep(0, 7),
                         y2011 = rep(0, 7),
                         y2012 = rep(0, 7),
                         y2013 = rep(0, 7),
                         y2014 = rep(0, 7),
                         y2015 = rep(0, 7),
                         y2016 = rep(0, 7),
                         y2017 = rep(0, 7),
                         y2018 = rep(0, 7),
                         y2019 = rep(0, 7),
                         y2020 = rep(0, 7))
  
  # Subsetting the merged dataset to the study of interest
  year_vec <- 2010:2020
  us_data_sub <- us_data[us_data$cStudy %in% studies, ]
  for (i in 1:length(year_vec)) {
    us_data_sub2 <- us_data_sub[us_data_sub$cEnrollmentYear == year_vec[i], ]
    
    # Counting the races
    race_counts <- plyr::count(us_data_sub2, 'cRace')
    race_counts$prop <- race_counts$freq / sum(race_counts$freq)
    
    # Adding missing races
    miss_race <- prop_res$Race[!(prop_res$Race %in% race_counts$cRace)]
    
    if (length(miss_race) != 0) {
      race_counts <- rbind(race_counts, data.frame(cRace = miss_race, 
                                                 freq =  rep(0,length(miss_race)),
                                                 prop = rep(0,length(miss_race))))
    }
    
    
    # reordering the dataframe 
    race_counts <- race_counts[order(race_counts$cRace), ]
    
  
  # Adding to dataframe
    prop_res[, i + 1] <- race_counts$prop
  }
  return (prop_res)
}
```

## Pipeline for calculating proportions over time - cumulative
```{r}
props_over_timeC <- function(studies) {
  
  # make dataframe to store results
  prop_res <- data.frame(Race = c('AA', 'AE', 'AS', 'CA', 'MR', 'NH', 'OT'),
                         y2010 = rep(0, 7),
                         y2011 = rep(0, 7),
                         y2012 = rep(0, 7),
                         y2013 = rep(0, 7),
                         y2014 = rep(0, 7),
                         y2015 = rep(0, 7),
                         y2016 = rep(0, 7),
                         y2017 = rep(0, 7),
                         y2018 = rep(0, 7),
                         y2019 = rep(0, 7),
                         y2020 = rep(0, 7))
  
  # Subsetting the merged dataset to the study of interest
  year_vec <- 2010:2020
  us_data_sub <- us_data[us_data$cStudy %in% studies, ]

  cumulative_count <- rep(0, 7)
  for (i in 1:length(year_vec)) {
    us_data_sub2 <- us_data_sub[us_data_sub$cEnrollmentYear <= year_vec[i], ]
    
    # Counting the races
    race_counts <- plyr::count(us_data_sub2, 'cRace')
    
    race_counts$prop <- race_counts$freq / sum(race_counts$freq)
    
    # Adding missing races
    miss_race <- prop_res$Race[!(prop_res$Race %in% race_counts$cRace)]
    
    if (length(miss_race) != 0) {
      race_counts <- rbind(race_counts, data.frame(cRace = miss_race, 
                                                 freq =  rep(0,length(miss_race)),
                                                 prop = rep(0,length(miss_race))))
    }
    
    
    # reordering the dataframe 
    race_counts <- race_counts[order(race_counts$cRace), ]
    
    cumulative_count <- cumulative_count + race_counts$freq
    names(cumulative_count) <- race_counts$cRace
    
  # Adding to dataframe
    prop_res[, i + 1] <- cumulative_count / max(c(sum(cumulative_count),1))
  }
  return (prop_res)
}
```

## Plots
```{r}
my_colors <- c('black', '#F8766D', '#85AD00', '#00A5FF', '#CF9400', '#FF689F', '#80796BFF')
```

### GPC: Proportions over Time
```{r}
pdf('GPC_plots.pdf')
plot(NA, xlim = c(2010,2020), ylim = c(0,1),
     xlab = 'Year', ylab = 'Proportion',
     main = 'GPC: Race Proportions over Time')
grid()

gpc_obj <- props_over_time('GPC')

for (i in 1:nrow(gpc_obj)) {
  lines(2010:2020, gpc_obj[i,2:ncol(gpc_obj)], col = my_colors[i], lwd = 1.4)
}

legend('top',
       legend = gpc_obj$Race,
       col = my_colors,
       lwd = 1.4,
       lty = 1, horiz = T,
       cex = 0.65)
```

### GPC: Cumulative Proportion over Time
```{r}
plot(NA, xlim = c(2010,2020), ylim = c(0,1),
     xlab = 'Year', ylab = 'Proportion',
     main = 'GPC: Cumulative Race Proportions over Time')
grid()

gpc_obj <- props_over_timeC('GPC')

for (i in 1:nrow(gpc_obj)) {
  lines(2010:2020, gpc_obj[i,2:ncol(gpc_obj)], col = my_colors[i], lwd = 1.4)
}

legend('top',
       legend = gpc_obj$Race,
       col = my_colors,
       lwd = 1.4,
       lty = 1, horiz = T,
       cex = 0.65)
dev.off()
```

### COGS2: Proportion over Time
```{r}
pdf('COGS2_plots.pdf')

plot(NA, xlim = c(2010,2020), ylim = c(0,1),
     xlab = 'Year', ylab = 'Proportion',
     main = 'COGS2: Race Proportions over Time')
grid()

cogs2_obj <- props_over_time('COGS2')

for (i in 1:nrow(cogs2_obj)) {
  lines(2010:2020, cogs2_obj[i,2:ncol(cogs2_obj)], col = my_colors[i], lwd = 1.4)
}

legend('top',
       legend = cogs2_obj$Race,
       col = my_colors,
       lwd = 1.4,
       lty = 1, horiz = T,
       cex = 0.65)
```

### COGS2: Cumulative Proportion over Time
```{r}
plot(NA, xlim = c(2010,2020), ylim = c(0,1),
     xlab = 'Year', ylab = 'Proportion',
     main = 'COGS2: Cumulative Race Proportions over Time')
grid()

cogs2_obj <- props_over_timeC('COGS2')

for (i in 1:nrow(cogs2_obj)) {
  lines(2010:2020, cogs2_obj[i,2:ncol(cogs2_obj)], col = my_colors[i], lwd = 1.4)
}

legend('top',
       legend = cogs2_obj$Race,
       col = my_colors,
       lwd = 1.4,
       lty = 1, horiz = T,
       cex = 0.65)

dev.off()
```

### BSNIP: Proportions over Time
```{r}
pdf('BSNIP_plots.pdf')
plot(NA, xlim = c(2010,2020), ylim = c(0,1),
     xlab = 'Year', ylab = 'Proportion',
     main = 'BSNIP: Race Proportions over Time')
grid()

bsnip_obj <- props_over_time(c('BSNIP1', 'BSNIP2'))

for (i in 1:nrow(bsnip_obj)) {
  lines(2010:2020, bsnip_obj[i,2:ncol(bsnip_obj)], col = my_colors[i], lwd = 1.4)
}

legend('top',
       legend = bsnip_obj$Race,
       col = my_colors,
       lwd = 1.4,
       lty = 1, horiz = T,
       cex = 0.65)


```

### BSNIP: Cumulative Proportion over Time
```{r}
plot(NA, xlim = c(2010,2020), ylim = c(0,1),
     xlab = 'Year', ylab = 'Proportion',
     main = 'BSNIP: Cumulative Race Proportions over Time')
grid()

bsnip_obj <- props_over_timeC(c('BSNIP1', 'BSNIP2'))

for (i in 1:nrow(bsnip_obj)) {
  lines(2010:2020, bsnip_obj[i,2:ncol(bsnip_obj)], col = my_colors[i], lwd = 1.4)
}

legend('top',
       legend = bsnip_obj$Race,
       col = my_colors,
       lwd = 1.4,
       lty = 1, horiz = T,
       cex = 0.65)

dev.off()
```

## Exploring BSNIP further: BSNIP1 and BSNIP2
### BSNIP1: Proportions over Time
```{r}
pdf('BSNIP12_over_time.pdf')
plot(NA, xlim = c(2010,2020), ylim = c(0,1),
     xlab = 'Year', ylab = 'Proportion',
     main = 'BSNIP1: Race Proportions over Time')
grid()

bsnip1_obj <- props_over_time('BSNIP1')

for (i in 1:nrow(bsnip1_obj)) {
  lines(2010:2020, bsnip1_obj[i,2:ncol(bsnip1_obj)], col = my_colors[i], lwd = 1.4)
}

legend('top',
       legend = bsnip1_obj$Race,
       col = my_colors,
       lwd = 1.4,
       lty = 1, horiz = T,
       cex = 0.65)
```

### BSNIP2: Proportion over Time
```{r}
plot(NA, xlim = c(2010,2020), ylim = c(0,1),
     xlab = 'Year', ylab = 'Proportion',
     main = 'BSNIP2: Race Proportions over Time')
grid()

bsnip2_obj <- props_over_time('BSNIP2')

for (i in 1:nrow(bsnip2_obj)) {
  lines(2010:2020, bsnip2_obj[i,2:ncol(bsnip2_obj)], col = my_colors[i], lwd = 1.4)
}

legend('top',
       legend = bsnip2_obj$Race,
       col = my_colors,
       lwd = 1.4,
       lty = 1, horiz = T,
       cex = 0.65)
dev.off()
```


### BSNIP1: Cumulative Proportion over Time
```{r}
pdf('BSNIP12_cumulative.pdf')
plot(NA, xlim = c(2010,2020), ylim = c(0,1),
     xlab = 'Year', ylab = 'Proportion',
     main = 'BSNIP1: Cumulative Race Proportions over Time')
grid()

bsnip1_obj <- props_over_timeC('BSNIP1')

for (i in 1:nrow(bsnip1_obj)) {
  lines(2010:2020, bsnip1_obj[i,2:ncol(bsnip1_obj)], col = my_colors[i], lwd = 1.4)
}

legend('top',
       legend = bsnip1_obj$Race,
       col = my_colors,
       lwd = 1.4,
       lty = 1, horiz = T,
       cex = 0.65)
```

### BSNIP2: Cumulative Proportion over Time
```{r}
plot(NA, xlim = c(2010,2020), ylim = c(0,1),
     xlab = 'Year', ylab = 'Proportion',
     main = 'BSNIP2: Cumulative Race Proportions over Time')
grid()

bsnip2_obj <- props_over_timeC('BSNIP2')

for (i in 1:nrow(bsnip2_obj)) {
  lines(2010:2020, bsnip2_obj[i,2:ncol(bsnip2_obj)], col = my_colors[i], lwd = 1.4)
}

legend('top',
       legend = bsnip2_obj$Race,
       col = my_colors,
       lwd = 1.4,
       lty = 1, horiz = T,
       cex = 0.65)

dev.off()
```

