---
title: "Variation in COGS2/BSNIP"
author: "Daniel Zoleikhaeian"
date: "2023-05-03"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading libraries
```{r}
library(plyr)
library(dplyr)
library(ggplot2)
library(readxl)
```

# Census Data

## Datasets
```{r}
dat_2010 <- read_xlsx('C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\reproj_30_2010.xlsx')
dat_2020 <- read_xlsx('C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\reproj_30_2020.xlsx')

dat_2010_races <- dat_2010[,2:8]
dat_2020_races <- dat_2020[,2:8]

rownames(dat_2010_races) <- dat_2010$City
rownames(dat_2020_races) <- dat_2020$City
#View(dat_2010_races)
```

## Making proportions
```{r}

# Making proportions
prop_2010 <- as.data.frame(t(apply(dat_2010_races, MARGIN = 1, function(x) {x / sum(x)}, simplify = T)))

# Making a city column
prop_2010$City <- rownames(prop_2010)
prop_2010 <- prop_2010[, c(ncol(prop_2010), 1:(ncol(prop_2010) - 1))]
rownames(prop_2010) <- NULL

# reformatting the City vector
prop_2010$City <- lapply(prop_2010$City, function (x)  {substr(x, 1, unlist(gregexpr(',', x))[1] - 1)}  )

# Adding a Year column
prop_2010$Year <- rep(2010, nrow(prop_2010))

prop_2020 <- as.data.frame(t(apply(dat_2020_races, MARGIN = 1, function(x) {x / sum(x)}, simplify = T)))

prop_2020$City <- rownames(prop_2020)
prop_2020 <- prop_2020[, c(ncol(prop_2020), 1:(ncol(prop_2020) - 1))]
rownames(prop_2020) <- NULL

prop_2020$City <- lapply(prop_2020$City, function (x)  {substr(x, 1, unlist(gregexpr(',', x))[1] - 1)}  )

prop_2020$Year <- rep(2020, nrow(prop_2020))

# stacking the dataframes on top of each other
prop_2010_2020 <- rbind(prop_2010, prop_2020)

# checking if the rows are equal
nrow(prop_2010_2020) == (nrow(prop_2010) + nrow(prop_2020))

# standardizing acronyms
colnames(prop_2010_2020)[colnames(prop_2010_2020) == 'NA'] <- 'AE'
colnames(prop_2010_2020)[colnames(prop_2010_2020) == 'PI'] <- 'NH'
```

## Making LM objects for AA within each cities in prop_2010_2020
```{r}
lm_objects <- vector("list", nrow(prop_2010))
cities <- prop_2010$City
for (i in 1:length(cities)) {
  lm_obj <- lm(AA ~ Year, data = prop_2010_2020[unlist(prop_2010_2020$City) == cities[i], ])
  lm_objects[[i]] <- lm_obj
}

unlist(prop_2010_2020$City)
names(lm_objects) <- cities
```

## Calculating the 2010-2013 expected proportion based on LM results
```{r}
year_in <- mean(c(2010, 2013))
year_exp_props <- rep(0, length(cities))
for (i in 1:length(year_exp_props)) {
  year_exp_props[i] <- predict(lm_objects[[i]], newdata = data.frame(Year = year_in))
}

names(year_exp_props) <- cities
year_exp_props # expected proportion of AA for 2010-2013 chunk across all cities
```

# Loading Dataset
```{r}
us_data <- read.csv('C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\bgc_merge_cDiag123_new.csv')
us_data$cEnrollmentYear <- as.numeric(substr(us_data$cEnrollmentDateYear, 1,4))
```

# Data Subsets by Year Range
```{r}
bsnip1_1013 <- us_data[us_data$cEnrollmentYear %in% 2010:2013 & us_data$cStudy == 'BSNIP1', ]
cogs2_1013 <- us_data[us_data$cEnrollmentYear %in% 2010:2013 & us_data$cStudy == 'COGS2', ]
gpc_1013 <- us_data[us_data$cEnrollmentYear %in% 2010:2013 & us_data$cStudy == 'GPC', ]

# just checking bsnip2 between 2010-2013; should be empty
bsnip2_1013 <- us_data[us_data$cEnrollmentYear %in% 2010:2013 & us_data$cStudy == 'BSNIP2', ] # 0 rows
```

# Getting Race Counts and Proportions
```{r}
# Race counts
race_1013_gpc <- plyr::count(gpc_1013, c('cLocationCity', 'cRace'))
race_1013_bsnip1 <- plyr::count(bsnip1_1013, c('cLocationCity', 'cRace'))
race_1013_cogs2 <- plyr::count(cogs2_1013, c('cLocationCity', 'cRace'))

city_reps_1013_gpc <- plyr::count(race_1013_gpc[,1:2], 'cLocationCity')
city_reps_1013_bsnip1 <- plyr::count(race_1013_bsnip1[,1:2], 'cLocationCity')
city_reps_1013_cogs2 <- plyr::count(race_1013_cogs2[,1:2], 'cLocationCity')

city_1013_gpc <- plyr::count(gpc_1013, 'cLocationCity')
city_1013_bsnip1 <- plyr::count(bsnip1_1013, 'cLocationCity')
city_1013_cogs2 <- plyr::count(cogs2_1013, 'cLocationCity')

tots_1013_gpc <- rep(city_1013_gpc$freq, city_reps_1013_gpc$freq)
tots_1013_bsnip1 <- rep(city_1013_bsnip1$freq, city_reps_1013_bsnip1$freq)
tots_1013_cogs2 <- rep(city_1013_cogs2$freq, city_reps_1013_cogs2$freq)


race_1013_gpc$prop <- race_1013_gpc$freq / tots_1013_gpc
race_1013_cogs2$prop <- race_1013_cogs2$freq / tots_1013_cogs2
race_1013_bsnip1$prop <- race_1013_bsnip1$freq / tots_1013_bsnip1

race_1013_gpc$Study <- rep('GPC', nrow(race_1013_gpc))
race_1013_bsnip1$Study <- rep('BSNIP1', nrow(race_1013_bsnip1))
race_1013_cogs2$Study <- rep('COGS2', nrow(race_1013_cogs2))

race_total <- rbind(race_1013_gpc, race_1013_cogs2, race_1013_bsnip1)
```

# Plotting counts and proportions

## Helper Function
```{r}
plot_Bar <- function(data_in, y_col, y_name, race_acr, fill_arg, yr_range) {
  ggplot(data = data_in, aes_string(x = 'cLocationCity', y = y_col, fill = fill_arg)) +
  geom_bar(stat = "identity", position='dodge') + 
  ggtitle(paste(race_acr,  y_name, yr_range))
}

## saving plots as pdf file
pdf('30aa_COGS2-BSNIP_Variation_1013.pdf')
```

## Plotting race proportions for the 3 studies
```{r}
## Just observing AA
plot_Bar(race_total[race_total$cRace == 'AA', ], 'prop', 'Proportions', race_acr = 'AA', fill_arg = 'Study', yr_range = '2010-2013') + coord_flip()

```


## Getting proportion deltas
```{r}
AA_sub <- race_total[race_total$cRace == 'AA', ]

# observed - expected
AA_sub$prop_delt <- rep(0, nrow(AA_sub))

## subset the AA_sub further by only looking at the census represented cities
AA_sub <- AA_sub[AA_sub$cLocationCity %in% cities, ]

# Loop through rows of AA_sub
for (i in 1:nrow(AA_sub)) {
  # Match the city from AA_sub with the city of the expected prop vector
  AA_sub$prop_delt[i] <- AA_sub$prop[i] - year_exp_props[names(year_exp_props) == AA_sub$cLocationCity[i]]
}
```

## Plotting Proportion Deltas
```{r}
plot_Bar(data_in = AA_sub, y_col = 'prop_delt', y_name = 'Proportion Delta', race_acr = 'AA', fill_arg = 'Study', yr_range = '2010-2013') + coord_flip()
```

## Characterizing the variability

Methods used:
* Range of proportion differences
* Mean and standard deviation
  * sensitive to outliers
* Median and IQR
  * robust to outliers

```{r}

GPC_mean <- mean(AA_sub$prop_delt[AA_sub$Study == 'GPC'])
GPC_sd <- sd(AA_sub$prop_delt[AA_sub$Study == 'GPC'])

COGS2_mean <- mean(AA_sub$prop_delt[AA_sub$Study == 'COGS2'])
COGS2_sd <- sd(AA_sub$prop_delt[AA_sub$Study == 'COGS2'])

BSNIP1_mean <- mean(AA_sub$prop_delt[AA_sub$Study == 'BSNIP1'])
BSNIP1_sd <- sd(AA_sub$prop_delt[AA_sub$Study == 'BSNIP1'])

prop_delts_1013 <- data.frame(means = c(GPC_mean, COGS2_mean, BSNIP1_mean), 
                              st_devs = c(GPC_sd, COGS2_sd, BSNIP1_sd), 
                              Study = c('GPC', 'COGS2', 'BSNIP1'))

bar_plot <- ggplot(data = prop_delts_1013, aes(x = Study, y = means, fill = Study)) +
  geom_bar(stat = "identity", position='dodge') +
  geom_errorbar(aes(ymin=means-st_devs, ymax=means+st_devs), width=.2,
                 position=position_dodge(.9)) + 
  labs(y = 'Mean Proportion Deltas') +
  ggtitle('2010-2013: Mean Proportion Deltas with Standard Deviations')

bar_plot + geom_point(data = AA_sub, aes(x = Study, y = prop_delt))
```

## Beeswarm version: Boxplot + raw points
```{r}
library(ggbeeswarm)
beeswarm.plot <- ggplot(data = AA_sub, aes(x = Study, y = prop_delt, color = Study)) +
  geom_beeswarm(cex = 3, size = 1.75) + 
  scale_color_brewer(palette = "Set1") +
  labs(y = 'Proportion Deltas') +
  ggtitle('2010-2013: Proportion Deltas with Boxplot Overlay') 

beeswarm.plot + geom_boxplot(data = AA_sub, aes(x = Study, y = prop_delt), alpha = 0.25)

# closing the pdf
dev.off()
```













