---
title: "GPC Variation, 2010-2020 actual years"
author: "Daniel Zoleikhaeian"
date: "2023-05-15"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading libraries
```{r}
library(plyr)
library(dplyr)
library(ggplot2)
library(readxl)
```

# Census Data

## Datasets
```{r}
dat_2010 <- read_xlsx('C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\reproj_30_2010.xlsx')
dat_2020 <- read_xlsx('C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\reproj_30_2020.xlsx')

dat_2010_races <- dat_2010[,2:8]
dat_2020_races <- dat_2020[,2:8]

rownames(dat_2010_races) <- dat_2010$City
rownames(dat_2020_races) <- dat_2020$City
#View(dat_2010_races)
```

## Making proportions
```{r}

# Making proportions
prop_2010 <- as.data.frame(t(apply(dat_2010_races, MARGIN = 1, function(x) {x / sum(x)}, simplify = T)))

# Making a city column
prop_2010$City <- rownames(prop_2010)
prop_2010 <- prop_2010[, c(ncol(prop_2010), 1:(ncol(prop_2010) - 1))]
rownames(prop_2010) <- NULL

# reformatting the City vector
prop_2010$City <- lapply(prop_2010$City, function (x)  {substr(x, 1, unlist(gregexpr(',', x))[1] - 1)}  )

# Adding a Year column
prop_2010$Year <- rep(2010, nrow(prop_2010))

prop_2020 <- as.data.frame(t(apply(dat_2020_races, MARGIN = 1, function(x) {x / sum(x)}, simplify = T)))

prop_2020$City <- rownames(prop_2020)
prop_2020 <- prop_2020[, c(ncol(prop_2020), 1:(ncol(prop_2020) - 1))]
rownames(prop_2020) <- NULL

prop_2020$City <- lapply(prop_2020$City, function (x)  {substr(x, 1, unlist(gregexpr(',', x))[1] - 1)}  )

prop_2020$Year <- rep(2020, nrow(prop_2020))

# stacking the dataframes on top of each other
prop_2010_2020 <- rbind(prop_2010, prop_2020)

# checking if the rows are equal
nrow(prop_2010_2020) == (nrow(prop_2010) + nrow(prop_2020))

# standardizing acronyms
colnames(prop_2010_2020)[colnames(prop_2010_2020) == 'NA'] <- 'AE'
colnames(prop_2010_2020)[colnames(prop_2010_2020) == 'PI'] <- 'NH'
```

## Making LM objects for AA within each cities in prop_2010_2020
```{r}
lm_objects <- vector("list", nrow(prop_2010))
cities <- prop_2010$City
for (i in 1:length(cities)) {
  lm_obj <- lm(AA ~ Year, data = prop_2010_2020[unlist(prop_2010_2020$City) == cities[i], ])
  lm_objects[[i]] <- lm_obj
}

names(lm_objects) <- cities
```

## Calculating the 2010-2013 expected proportions based on LM results
```{r}
# check years 2010-2013
years_in <- 2010:2020

# dataframe structure
exp_props <- data.frame(City = unlist(cities), 
                        y2010 = rep(0, length(cities)),
                        y2011 = rep(0, length(cities)),
                        y2012 = rep(0, length(cities)),
                        y2013 = rep(0, length(cities)),
                        y2014 = rep(0, length(cities)),
                        y2015 = rep(0, length(cities)),
                        y2016 = rep(0, length(cities)),
                        y2017 = rep(0, length(cities)),
                        y2018 = rep(0, length(cities)),
                        y2019 = rep(0, length(cities)),
                        y2020 = rep(0, length(cities)))

# iterate through all the cities, then fill in the data
for (i in 1:nrow(exp_props)) {
  exp_props[i, 2:ncol(exp_props)] <- predict(lm_objects[[i]], newdata = data.frame(Year = years_in))
}

exp_props # expected proportion of AA for 2010-2013 chunk across all cities

# transform to long format
long_exp_props <- data.frame(City = rep(unlist(cities)),
                             AA_prop = c(exp_props$y2010,
                                         exp_props$y2011,
                                         exp_props$y2012,
                                         exp_props$y2013,
                                         exp_props$y2014,
                                         exp_props$y2015,
                                         exp_props$y2016,
                                         exp_props$y2017,
                                         exp_props$y2018,
                                         exp_props$y2019,
                                         exp_props$y2020),
                             Year = rep(2010:2020, rep(length(unlist(cities)), length(years_in)))
                             )
```

# Loading Dataset
```{r}
us_data <- read.csv('C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\bgc_merge_cDiag123_new.csv')
us_data$cEnrollmentYear <- as.numeric(substr(us_data$cEnrollmentDateYear, 1,4))
```

# Data Subsets by Year Range
```{r}

gpc_data <- us_data[us_data$cStudy == 'GPC', ]

gpc_1013 <- gpc_data[gpc_data$cEnrollmentYear %in% 2010:2013, ]
gpc_1416 <- gpc_data[gpc_data$cEnrollmentYear %in% 2014:2016, ]
gpc_1720 <- gpc_data[gpc_data$cEnrollmentYear %in% 2017:2020, ]

gpc_1020 <- list(gpc_1013, gpc_1416, gpc_1720)
```

# Getting Race Counts and Proportions from the studies
```{r}

# studies
year_chunks <- c('2010-2013', '2014-2016', '2017-2020')

# list to store dataframes
## Structure:
### Level 1: Year chunks
### Level 2: Years
### Level 3: dataframe with cities
chunk_props <- vector("list", 3)
names(chunk_props) <- year_chunks

# helper function: race counts to proportion
count2prop <- function(df_in, year) {
  
  df_sub <- df_in[df_in$cEnrollmentYear == year, ]
  race_ct <- plyr::count(df_sub, c('cLocationCity', 'cRace'))
  city_reps <- plyr::count(race_ct[,1:2], 'cLocationCity')
  city_totals <- plyr::count(df_sub, 'cLocationCity')
  tots_repped <- rep(city_totals$freq, city_reps$freq)
  race_ct$prop <- race_ct$freq / tots_repped
  return(race_ct)
  
}

temp_yr_df_list <- vector("list", length(years_in))
names(temp_yr_df_list) <- 2010:2020

# level 1
for (i in 1:length(chunk_props)) {
  # level 2
  for (j in 1:length(years_in)) {
    # level 3
    temp_yr_df_list[[j]] <- count2prop(gpc_1020[[i]], years_in[j])
    temp_yr_df_list[[j]]$Year_range <- rep(year_chunks[i], nrow(temp_yr_df_list[[j]]))
    temp_yr_df_list[[j]]$Year <- rep(years_in[j], nrow(temp_yr_df_list[[j]]))
  }
  chunk_props[[i]] <- temp_yr_df_list
}

# checking object
chunk_props ## as expected

# Merging all the dataframes into one object

chunk_props_unlist_inner <- list(dplyr::bind_rows(chunk_props[[1]], .id='column_label'), dplyr::bind_rows(chunk_props[[2]], .id='column_label'), dplyr::bind_rows(chunk_props[[3]], .id='column_label'))  

chunk_props_concat <- dplyr::bind_rows(chunk_props_unlist_inner, .id='column_label')

```

# Plotting counts and proportions

## Additional Helper Function
```{r}
plot_Lines <- function(data_in, ycol, yname, yr_range, study) {
  data_in <- data_in[data_in$Year_range == yr_range, ]
  ggplot(data_in, aes_string(x='Year',
                      y=ycol, col = 'cLocationCity'
                     )) +
  geom_point(cex = 1.5) + 
  geom_line(lwd = 1) +
  ylab(yname) +
  ggtitle(paste(study, ': ', ' AA ', yname, ' ', yr_range, sep = ''))
  
}
```

## Plotting race proportions for the 3 studies

```{r}

## saving plots as pdf file
pdf('C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\Rmd Files\\AA Proportion Variation over Time\\30bb_GPC_Variation_1020_actual_year.pdf')
    
    
## Just observing AA
AA_sub <- chunk_props_concat[chunk_props_concat$cRace == 'AA', ]

plot_Lines(data_in = AA_sub, yr_range = "2010-2013", ycol = 'prop', yname= 'Proportion', study = "GPC")
plot_Lines(data_in = AA_sub, yr_range = "2014-2016", ycol = 'prop', yname= 'Proportion', study = "GPC")
plot_Lines(data_in = AA_sub, yr_range = "2017-2020", ycol = 'prop', yname= 'Proportion', study = "GPC")
```

## Getting proportion deltas
```{r}

## subset the AA_sub further by only looking at the census represented cities
AA_sub <- AA_sub[AA_sub$cLocationCity %in% cities, ]

# observed - expected
AA_sub$prop_delt <- rep(0, nrow(AA_sub))

# Loop through rows of AA_sub
for (i in 1:nrow(AA_sub)) {
  # Match the city and year from AA_sub with the city and year of the expected prop dataframe
  AA_sub$prop_delt[i] <- AA_sub$prop[i] - long_exp_props$AA_prop[long_exp_props$City == AA_sub$cLocationCity[i] & long_exp_props$Year == AA_sub$Year[i]]
}
```

## Plotting Proportion Deltas by Year: FIX this
Idea:
* Make scatter plot (1 for each study)
- lines for each city
- x-axis would be year
- y-axis would be proportion delta
```{r}
plot_Lines(data_in = AA_sub, yr_range = "2010-2013", ycol = 'prop_delt', yname= 'Proportion Delta (Observed- Expected)', study = "GPC")
plot_Lines(data_in = AA_sub, yr_range = "2014-2016", ycol = 'prop_delt', yname= 'Proportion Delta (Observed- Expected)', study = "GPC")
plot_Lines(data_in = AA_sub, yr_range = "2017-2020", ycol = 'prop_delt', yname= 'Proportion Delta (Observed- Expected)', study = "GPC")
```

## Characterizing the variability: 

Methods used:
* Range of proportion differences
* Mean and standard deviation
  * sensitive to outliers
* Median and IQR
  * robust to outliers

```{r}

GPC10_mean <- mean(AA_sub$prop_delt[AA_sub$Year_range == '2010-2013'])
GPC10_sd <- sd(AA_sub$prop_delt[AA_sub$Year_range == '2010-2013'])

GPC14_mean <- mean(AA_sub$prop_delt[AA_sub$Year_range == '2014-2016'])
GPC14_sd <- sd(AA_sub$prop_delt[AA_sub$Year_range == '2014-2016'])

GPC17_mean <- mean(AA_sub$prop_delt[AA_sub$Year_range == '2017-2020'])
GPC17_sd <- sd(AA_sub$prop_delt[AA_sub$Year_range == '2017-2020'])

prop_delts_1020 <- data.frame(means = c(GPC10_mean, GPC14_mean, GPC17_mean), 
                              st_devs = c(GPC10_sd, GPC14_sd, GPC17_sd), 
                              Year_range = c('2010-2013', '2014-2016', '2017-2020'))

bar_plot <- ggplot(data = prop_delts_1020, aes(x = Year_range, y = means, fill = Year_range)) +
  geom_bar(stat = "identity", position='dodge') +
  geom_errorbar(aes(ymin=means-st_devs, ymax=means+st_devs), width=.2,
                 position=position_dodge(.9)) + 
  labs(y = 'Mean Proportion Deltas') +
  ggtitle('2010-2020: GPC AA Mean Proportion Deltas with Standard Deviations')

bar_plot + geom_point(data = AA_sub, aes(x = Year_range, y = prop_delt))
```

## Beeswarm version: Boxplot + raw points
```{r}
library(ggbeeswarm)
beeswarm.plot <- ggplot(data = AA_sub, aes(x = Year_range, y = prop_delt, color = Year_range)) +
  geom_beeswarm(cex = 3, size = 1.75) + 
  scale_color_brewer(palette = "Set1") +
  labs(y = 'Proportion Deltas') +
  ggtitle('2010-2020: GPC AA Proportion Deltas with Boxplot Overlay') 

beeswarm.plot + geom_boxplot(data = AA_sub, aes(x = Year_range, y = prop_delt), alpha = 0.25)

# closing the pdf
dev.off()
```