---
title: "32. COGS2 Proportion Deviations Setup"
author: "Daniel Zoleikhaeian"
date: "2023-05-22"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Libraries
```{r}
library(plyr)
library(dplyr)
library(ggplot2)
library(readxl)
```

# Census Data and Proportions

## Datasets
```{r}
dat_2010 <- read_xlsx('C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\reproj_30_2010.xlsx')
dat_2020 <- read_xlsx('C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\reproj_30_2020.xlsx')

dat_2010_races <- dat_2010[,2:8]
dat_2020_races <- dat_2020[,2:8]

sum(dat_2010[dat_2010$City == 'Los Angeles, CA', 2:8])
dat_2010

rownames(dat_2010_races) <- dat_2010$City
rownames(dat_2020_races) <- dat_2020$City
#View(dat_2010_races)
```

## Making proportions
```{r}

# Making proportions
prop_2010 <- as.data.frame(t(apply(dat_2010_races, MARGIN = 1, function(x) {x / sum(x)}, simplify = T)))

# Making a city column
prop_2010$City <- rownames(prop_2010)
prop_2010 <- prop_2010[, c(ncol(prop_2010), 1:(ncol(prop_2010) - 1))]
rownames(prop_2010) <- NULL

# reformatting the City vector
prop_2010$City <- lapply(prop_2010$City, function (x)  {substr(x, 1, unlist(gregexpr(',', x))[1] - 1)}  )

# Adding a Year column
prop_2010$Year <- rep(2010, nrow(prop_2010))

prop_2020 <- as.data.frame(t(apply(dat_2020_races, MARGIN = 1, function(x) {x / sum(x)}, simplify = T)))

prop_2020$City <- rownames(prop_2020)
prop_2020 <- prop_2020[, c(ncol(prop_2020), 1:(ncol(prop_2020) - 1))]
rownames(prop_2020) <- NULL

prop_2020$City <- lapply(prop_2020$City, function (x)  {substr(x, 1, unlist(gregexpr(',', x))[1] - 1)}  )

prop_2020$Year <- rep(2020, nrow(prop_2020))

# stacking the dataframes on top of each other
prop_2010_2020 <- rbind(prop_2010, prop_2020)

# checking if the rows are equal
nrow(prop_2010_2020) == (nrow(prop_2010) + nrow(prop_2020))

# standardizing acronyms
colnames(prop_2010_2020)[colnames(prop_2010_2020) == 'NA'] <- 'AE'
colnames(prop_2010_2020)[colnames(prop_2010_2020) == 'PI'] <- 'NH'
```

## Loading Dataset
```{r}
us_data <- read.csv('C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\bgc_merge_cDiag123_new.csv')
head(us_data)
us_data$cEnrollmentYear <- substr(us_data$cEnrollmentDateYear, 1,4)

# Combining OT/UNK and UNK into OT/UNK
us_data$cRace[us_data$cRace %in% c('UNK', 'OT/UNK')] <- 'OT'
cogs2 <- us_data[us_data$cStudy == 'COGS2', ]

```

## Pipeline for calculating proportions over time - not cumulative
* Modified from previous to account for cities
```{r}
props_over_time <- function(studies, city) {
  
  # make dataframe to store results
  prop_res <- data.frame(Race = c('AA', 'AE', 'AS', 'CA', 'MR', 'NH', 'OT'),
                         y2010 = rep(0, 7),
                         y2011 = rep(0, 7),
                         y2012 = rep(0, 7),
                         y2013 = rep(0, 7),
                         y2014 = rep(0, 7),
                         y2015 = rep(0, 7),
                         y2016 = rep(0, 7),
                         y2017 = rep(0, 7),
                         y2018 = rep(0, 7),
                         y2019 = rep(0, 7),
                         y2020 = rep(0, 7))
  
  # Subsetting the merged dataset to the study of interest
  year_vec <- 2010:2020
  us_data_sub <- us_data[us_data$cStudy %in% studies & us_data$cLocationCity == city, ]
  for (i in 1:length(year_vec)) {
    us_data_sub2 <- us_data_sub[us_data_sub$cEnrollmentYear == year_vec[i], ]
    
    # Counting the races
    race_counts <- plyr::count(us_data_sub2, 'cRace')
    race_counts$prop <- race_counts$freq / sum(race_counts$freq)
    
    # Adding missing races
    miss_race <- prop_res$Race[!(prop_res$Race %in% race_counts$cRace)]
    
    if (length(miss_race) != 0) {
      race_counts <- rbind(race_counts, data.frame(cRace = miss_race, 
                                                 freq =  rep(0,length(miss_race)),
                                                 prop = rep(0,length(miss_race))))
    }
    
    
    # reordering the dataframe 
    race_counts <- race_counts[order(race_counts$cRace), ]
    
  
  # Adding to dataframe
    prop_res[, i + 1] <- race_counts$prop
  }
  return (prop_res)
}
```

## Pipeline for calculating proportions over time - cumulative
* Modified from previous to account for cities
```{r}
props_over_timeC <- function(studies, city) {
  
  # make dataframe to store results
  prop_res <- data.frame(Race = c('AA', 'AE', 'AS', 'CA', 'MR', 'NH', 'OT'),
                         y2010 = rep(0, 7),
                         y2011 = rep(0, 7),
                         y2012 = rep(0, 7),
                         y2013 = rep(0, 7),
                         y2014 = rep(0, 7),
                         y2015 = rep(0, 7),
                         y2016 = rep(0, 7),
                         y2017 = rep(0, 7),
                         y2018 = rep(0, 7),
                         y2019 = rep(0, 7),
                         y2020 = rep(0, 7))
  
  # Subsetting the merged dataset to the study and city of interest
  year_vec <- 2010:2020
  us_data_sub <- us_data[us_data$cStudy %in% studies & us_data$cLocationCity == city, ]

  cumulative_count <- rep(0, 7)
  for (i in 1:length(year_vec)) {
    us_data_sub2 <- us_data_sub[us_data_sub$cEnrollmentYear <= year_vec[i], ]
    
    # Counting the races
    race_counts <- plyr::count(us_data_sub2, 'cRace')
    
    race_counts$prop <- race_counts$freq / sum(race_counts$freq)
    
    # Adding missing races
    miss_race <- prop_res$Race[!(prop_res$Race %in% race_counts$cRace)]
    
    if (length(miss_race) != 0) {
      race_counts <- rbind(race_counts, data.frame(cRace = miss_race, 
                                                 freq =  rep(0,length(miss_race)),
                                                 prop = rep(0,length(miss_race))))
    }
    
    
    # reordering the dataframe 
    race_counts <- race_counts[order(race_counts$cRace), ]
    
    cumulative_count <- cumulative_count + race_counts$freq
    names(cumulative_count) <- race_counts$cRace
    
  # Adding to dataframe
    prop_res[, i + 1] <- cumulative_count / max(c(sum(cumulative_count),1))
  }
  return (prop_res)
}
```

## Function to help plot cities
```{r}
plot_studyxcity <- function(fun_obj, study, city, col_pal, cum_spec = '') {
  plot(NA, xlim = c(2010,2015), ylim = c(0,1),
     xlab = 'Year', ylab = 'Proportion',
     main = paste(study, ' ',city, ': ', cum_spec, ' ', 'Race Proportions over Time', sep = ''))
  grid()
  plot_obj <- fun_obj(study, city)
  
  for (i in 1:nrow(plot_obj)) {
    lines(2010:2020, plot_obj[i,2:ncol(plot_obj)], col = col_pal[i], lwd = 1.4)
  }

  legend('top',
         legend = plot_obj$Race,
         col = col_pal,
         lwd = 1.4,
         lty = 1, horiz = T,
         cex = 0.65)
}
```

## COGS2: Proportion over Time- total
```{r}
my_colors <- c('black', '#F8766D', '#85AD00', '#00A5FF', '#CF9400', '#FF689F', '#80796BFF')


plot(NA, xlim = c(2010,2015), ylim = c(0,1),
     xlab = 'Year', ylab = 'Proportion',
     main = 'COGS2: Race Proportions over Time')
grid()


cogs2_obj <- props_over_time('COGS2')

for (i in 1:nrow(cogs2_obj)) {
  lines(2010:2020, cogs2_obj[i,2:ncol(cogs2_obj)], col = my_colors[i], lwd = 1.4)
}

legend('top',
       legend = cogs2_obj$Race,
       col = my_colors,
       lwd = 1.4,
       lty = 1, horiz = T,
       cex = 0.65)
```

### COGS2: Proportion over Time by City
```{r}

cogs2_cities <- unique(cogs2$cLocationCity)
for (i in 1:length(cogs2_cities)) {
  plot_studyxcity(props_over_time, 'COGS2', cogs2_cities[i], my_colors)
}

```

## COGS2: Cumulative Proportion over Time
```{r}
plot(NA, xlim = c(2010,2020), ylim = c(0,1),
     xlab = 'Year', ylab = 'Proportion',
     main = 'COGS2: Cumulative Race Proportions over Time')
grid()

cogs2_obj <- props_over_timeC('COGS2')

for (i in 1:nrow(cogs2_obj)) {
  lines(2010:2020, cogs2_obj[i,2:ncol(cogs2_obj)], col = my_colors[i], lwd = 1.4)
}

legend('top',
       legend = cogs2_obj$Race,
       col = my_colors,
       lwd = 1.4,
       lty = 1, horiz = T,
       cex = 0.65)


```

### COGS2:Cumulative Proportion over Time by City
```{r}
for (i in 1:length(cogs2_cities)) {
  plot_studyxcity(props_over_timeC, 'COGS2', cogs2_cities[i], my_colors, cum_spec='Cumulative')
}
```

## Generating Observed-Expected Information

### Making lm objects for each cogs2 city and each race
```{r}

## subsetting the proportions dataframe for only the cogs2 cities
cogs_10_20 <- prop_2010_2020[prop_2010_2020$City %in% cogs2_cities, ]

races <- sort(unique(cogs2$cRace))

lm_objects <- vector("list", length(cogs2_cities))
for (i in 1:length(lm_objects)) {
  em_list <- vector("list", length(races))
  names(em_list) <- races
  lm_objects[[i]] <- em_list
}
lm_objects

names(lm_objects) <- cogs2_cities
  

for (i in 1:length(cogs2_cities)) {
  for (j in 1:length(races)) {
    lm_obj <- lm(as.formula(paste0(races[j], "~ Year")), data = cogs_10_20[cogs_10_20$City == cogs2_cities[i], ])
    lm_objects[[i]][[j]] <- lm_obj
  }
}

# Setting all the names
# names(lm_objects$`San Diego`) <- races
# names(lm_objects$`Los Angeles`) <- races
# names(lm_objects$`Philadelphia`) <- races
# names(lm_objects$`Seattle`) <- races
# names(lm_objects$`New York`) <- races
lm_objects

```

### Calculating the 2010-2014 expected proportions based on LM results
```{r}

# empty list to hold all the dataframes for each city
exp_props_list <- vector(mode='list', length = length(cogs2_cities))
names(exp_props_list) <- cogs2_cities
exp_props_list


# check years 2010-2014
years_in <- 2010:2014


for(i in 1:length(cogs2_cities)) {
  # dataframe structure
  exp_props <- data.frame(Race = races, 
                        y2010 = rep(0, length(races)),
                        y2011 = rep(0, length(races)),
                        y2012 = rep(0, length(races)),
                        y2013 = rep(0, length(races)),
                        y2014 = rep(0, length(races)))
  # iterate through all the races, then fill in the data
  for (j in 1:nrow(exp_props)) {
    exp_props[j, 2:ncol(exp_props)] <- predict(lm_objects[[i]][[j]], newdata = data.frame(Year = years_in))
    
  # then place in the list
  }
  exp_props_list[[i]] <- exp_props
  
}
exp_props_list
```

### Helper function: Diversity index
Source: https://www2.census.gov/programs-surveys/demo/about/housing-patterns/multigroup_entropy.pdf

```{r}
mult_ent <- function(race_prop_vec) {
  
  tot <- 0
  
  for (i in 1:length(race_prop_vec)) {
    if (race_prop_vec[i] != 0) {
      tot <- tot + race_prop_vec[i] * log(1/race_prop_vec[i])
    }
  }
  return(tot)
}
```

### Generating the diversity index scores
```{r}

div <- c(Race = 'Diversity Index')

div_idxs <- data.frame(City = cogs2_cities,
                       y2010 = rep(0,length(cogs2_cities)),
                       y2011 = rep(0, length(cogs2_cities)),
                       y2012 = rep(0, length(cogs2_cities)),
                       y2013 = rep(0, length(cogs2_cities)),
                       y2014 = rep(0, length(cogs2_cities)))

for (i in 1:length(exp_props_list)) {
  div_idxs[i, 2:ncol(div_idxs)] <- apply(exp_props_list[[i]][,2:ncol(exp_props_list[[i]])], MARGIN = 2, mult_ent)
  rbind_vec <- c(div,apply(exp_props_list[[i]][,2:ncol(exp_props_list[[i]])], MARGIN = 2, mult_ent))
  exp_props_list[[i]] <- rbind(exp_props_list[[i]], rbind_vec)
}

## Table of Diversity Indices
div_idxs

write.csv(div_idxs, 'C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\diversity_idxs_cogs2.csv', row.names = F)
```




```{r}

```