---
title: "34. Replicating Diversity Index on National Equity Atlas"
author: "Daniel Zoleikhaeian"
date: "2023-05-31"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Loading COGS2 cities
```{r}
library(dplyr)
## 2010 IPUMS data
df_ipums <- read.csv('C:\\Users\\danie\\Documents\\Joshi Lab Materials\\usa_00002.csv')
df_COGS2 <- read.csv('C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\bgc_merge_cDiag123_new.csv')

df_COGS2 <- df_COGS2[df_COGS2$cStudy == 'COGS2', ]
sort(unique(df_COGS2$cLocationCity))

## Subsetting by COGS2 cities
### Using the city codes provided in the data dictionary
cogs2_citycodes <- c(3730, 4610, 5330, 6270, 6430)
df_ipums_cogs2 <- df_ipums[df_ipums$CITY %in% cogs2_citycodes, ]
unique(df_ipums_cogs2$CITY)
sort(unique(df_ipums$CITY))

# seems like no 2010 for San Diego from IPUMS
sum(df$CITY == 6270)

head(df_ipums_cogs2)

# removing people who did not report HISPAN status
df_9 <- df_ipums_cogs2[df_ipums_cogs2$HISPAN == 9, ] # empty dataframe
View(df_9)

# everyone reported a Hispanic status in the sample
unique(df_ipums_cogs2$HISPAN)
unique(df_ipums$HISPAN)
```

# Checking Los Angeles in 2010
```{r}
LA <- df_ipums_cogs2[df_ipums_cogs2$CITY == 3730, ]

# summing the PERWT column, and multiplying by 100 to check if this matches the population

sum(LA$PERWT) # 3795389
LA$CITYPOP[1] * 100 # 3797100 --> some discrepancy

head(LA)

# checking if all cases are complete (no NAs or blanks)
all(complete.cases(LA)) # returns TRUE

```

# Generating Diversity Index

* Source for what counts as PI: 

https://api-gbv.org/wp-content/uploads/2019/06/API-demographics-identities-May-2019.pdf

PI ethnicities: Carolinian, Chamorro, Chuukese, Fijian, Guamanian, Hawaiian, Kosraean,
Marshallese, Native Hawaiian, Niuean, Palauan, Papua New Guinean, Pohnpeian, Samoan, Tokelauan,
Tongan, Yapese.

```{r}
# Examining the Asian population
# goal is to find a way to separate other Asian and PI
# approach: look at RACED variable

other_as_pi <- LA[LA$RACE == 6, ]
sort(unique(other_as_pi$RACED))

# using data dictionary for RACED
PI_raced <- c(680:699)

PI_only <- LA[LA$RACE == 6 & LA$RACED %in% PI_raced, ]
AS_only <- LA[LA$RACE == 6 & !(LA$RACED %in% PI_raced), ]

(nrow(PI_only) + nrow(AS_only)) == nrow(other_as_pi)
```

## Encoding a new race/ethnicity variable
```{r}
# Using same 7 categories as National Equity Atlas
LA$NEA_grp <- rep(0, nrow(LA))
LA$NEA_grp[LA$RACE == 1 & LA$HISPAN == 0] <- 1 # Non-hispanic White
LA$NEA_grp[LA$RACE == 2 & LA$HISPAN == 0] <- 2 # Non-hispanic Black
LA$NEA_grp[LA$RACE == 3 & LA$HISPAN == 0] <- 3 # Non-hispanic American Indian or Alaska Native
LA$NEA_grp[LA$RACE %in% 4:6 & !(LA$RACED %in% PI_raced) & LA$HISPAN == 0] <- 4 # Non-hispanic Asian
LA$NEA_grp[LA$RACE == 6 & LA$RACED %in% PI_raced & LA$HISPAN == 0] <- 5 # Non-Hispanic Pacific Islander
LA$NEA_grp[LA$RACE %in% 7:9 & LA$HISPAN == 0] <- 6 # Non-Hispanic Mixed/Other
LA$NEA_grp[LA$HISPAN != 0] <- 7 # Hispanic
```

## Diversity index calculation

### Helper function: Diversity Index
```{r}
mult_ent <- function(race_prop_vec) {
  
  tot <- 0
  
  for (i in 1:length(race_prop_vec)) {
    if (race_prop_vec[i] != 0) {
      tot <- tot + race_prop_vec[i] * log(1/race_prop_vec[i])
    }
  }
  return(tot)
}
```

### Generating the vector of group proportions
```{r}
# effective total population
tot_pop <- sum(LA$PERWT)

# Generating a proportion
LA$PROP <- LA$PERWT/tot_pop
sum(LA$PROP) # sums to 1

LA$PROP2 <- LA$PERWT/(LA$CITYPOP*100)
sum(LA$PROP2) # nearly sums to 1

# Generating a new data vector for all the race categories from NEA_grp

prop_vec <- rep(0, 7)
prop_vec2 <- prop_vec

for (i in 1:length(prop_vec)) {
  prop_vec[i] <- sum(LA$PROP[LA$NEA_grp == i])
  prop_vec2[i] <- sum(LA$PROP2[LA$NEA_grp == i])
}

## calculating LA multigroup entropy aka diversity index in 2010
mult_ent(prop_vec) # 1.270208
mult_ent(prop_vec2) # 1.270086

## compare to NEA in LA 2010: 1.28

## same thing but combining Asian and Pacific Islander
AS_PI <- prop_vec[4] + prop_vec[5]
prop_vec3 <- c(prop_vec[1:3], AS_PI, prop_vec[6:7])
mult_ent(prop_vec3) # 1.263625
```

# Checking all COGS cities except SD
```{r}

cities <- unique(df_ipums_cogs2$CITY)
df_res <- data.frame(CityID = cities, 
                     City_name = c('Los Angeles', 'New York', 'Philadelphia', 'Seattle'),
                     Diversity_idx = rep(0, length(cities)))

# Using same 7 categories as National Equity Atlas
df_ipums_cogs2$NEA_grp <- rep(0, nrow(df_ipums_cogs2))
df_ipums_cogs2$NEA_grp[df_ipums_cogs2$RACE == 1 & df_ipums_cogs2$HISPAN == 0] <- 1 # Non-hispanic White
df_ipums_cogs2$NEA_grp[df_ipums_cogs2$RACE == 2 & df_ipums_cogs2$HISPAN == 0] <- 2 # Non-hispanic Black
df_ipums_cogs2$NEA_grp[df_ipums_cogs2$RACE == 3 & df_ipums_cogs2$HISPAN == 0] <- 3 # Non-hispanic American Indian or Alaska Native
df_ipums_cogs2$NEA_grp[df_ipums_cogs2$RACE %in% 4:6 & !(df_ipums_cogs2$RACED %in% PI_raced) & df_ipums_cogs2$HISPAN == 0] <- 4 # Non-hispanic Asian
df_ipums_cogs2$NEA_grp[df_ipums_cogs2$RACE == 6 & df_ipums_cogs2$RACED %in% PI_raced & df_ipums_cogs2$HISPAN == 0] <- 5 # Non-Hispanic Pacific Islander
df_ipums_cogs2$NEA_grp[df_ipums_cogs2$RACE %in% 7:9 & df_ipums_cogs2$HISPAN == 0] <- 6 # Non-Hispanic Mixed/Other
df_ipums_cogs2$NEA_grp[df_ipums_cogs2$HISPAN != 0] <- 7 # Hispanic

for (i in 1:length(cities)) {
  # subset to city of interest
  df_sub <- df_ipums_cogs2[df_ipums_cogs2$CITY == cities[i],]
  
  # using the city population for that city to generate a proportion
  df_sub$PROP <- df_sub$PERWT/(df_sub$CITYPOP*100)
  
  # generating proportion vector
  prop_vec <- rep(0, 7)
  for (j in 1:length(prop_vec)) {
    prop_vec[j] <- sum(df_sub$PROP[df_sub$NEA_grp == j])
  }
  
  # calculting multigroup entropy
  df_res$Diversity_idx[i] <- mult_ent(prop_vec)  
  
}

df_res$NEA_Diversity_idx <- c(1.28,1.42,1.24,1.10)

# displaying results
View(df_res)


```

