---
title: "Diversity Scores for HC and SZSAFD"
author: "Daniel Zoleikhaeian"
date: "2023-06-02"
output: pdf_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Importing data
```{r}
library(plyr)
library(dplyr)
df_COGS2 <- read.csv('C:\\Users\\danie\\Documents\\Joshi Lab Materials\\3 Studies Dataset\\Dataset Merge\\bgc_merge_cDiag123_060723.csv')
df_COGS2 <- df_COGS2[df_COGS2$cStudy == 'COGS2', ]
nrow(df_COGS2)

head(df_COGS2)
# adding year column
df_COGS2$cEnrollmentYear <- as.numeric(substr(df_COGS2$cEnrollmentDateYear, 1,4))
head(df_COGS2)

unique(df_COGS2$cRace)

# re-encoding OT/UNK/MR into one group
df_COGS2$cRace2 <- df_COGS2$cRace
df_COGS2$cRace2[df_COGS2$cRace2 %in% c('OT', 'OT/UNK', 'MR')] <- 'OT/MR'
nrow(df_COGS2)

# omitting NA
#df_COGS2 <- na.omit(df_COGS2)
nrow(df_COGS2)

df_COGS2_inc_cases <- df_COGS2[!complete.cases(df_COGS2),]
View(df_COGS2_inc_cases)

# email Joyce about COGS2 dates - DONE 6/7/23

```

### Helper function: Diversity Index
```{r}
mult_ent <- function(race_prop_vec) {
  
  tot <- 0
  
  for (i in 1:length(race_prop_vec)) {
    if (race_prop_vec[i] != 0) {
      tot <- tot + race_prop_vec[i] * log(1/race_prop_vec[i])
    }
  }
  return(tot)
}
```

## Generating Diversity Scores (Cumulative) by Year and City
```{r}
diagnoses <- c('CS', 'SZSAFD')
cities <- unique(df_COGS2$cLocationCity)
years <- unique(df_COGS2$cEnrollmentYear)

n <- 2 * length(cities) * length(years)
n2 <- 2 * length(cities)

# empty dataframe to hold data
df_di <- data.frame(City = rep('',n),
                    Year = rep(0,n),
                    Diagnosis = rep('',n),
                    mult_ent = rep(0,n)
                    )

df_di2 <- data.frame(City = rep('',n2),
                    Diagnosis = rep('',n2),
                    mult_ent = rep(0,n2)
                    )

# tertiary level: diagnosis
# outer level: year
# inner level: city
n_track <- 1
n_track2 <- 1
for (d in 1:length(diagnoses)) {
  for (i in 1:length(years)) {
    for (j in 1:length(cities)) {
      df_sub <- df_COGS2[df_COGS2$cDiagnosis3 == diagnoses[d] & df_COGS2$cLocationCity == cities[j] & df_COGS2$cEnrollmentYear <= years[i],]
      
      # if the dataframe is empty, skip this iteration
      if (nrow(df_sub) == 0) {
        df_di[n_track,1] <- cities[j]
        df_di[n_track,3] <- diagnoses[d]
        df_di_[n_track, 2] <- years[i]
        df_di[n_track,4] <- -99 # the code for no data
        n_track <- n_track + 1
        next
      }
      
      # first subset out the hispanics, then count up and make proportions out of the other races
      his_prop <- sum(df_sub$cHispanicorLatino == 'Yes')/nrow(df_sub)
      race_ct <- plyr::count(df_sub[df_sub$cHispanicorLatino == 'No',], 'cRace2')
      race_ct$prop <- race_ct$freq/nrow(df_sub)
      
      prop_vec <- c(his_prop, race_ct$prop)
      m_ent <- mult_ent(prop_vec)
      
      df_di[n_track,1] <- cities[j]
      df_di[n_track,2] <- years[i]
      df_di[n_track,3] <- diagnoses[d]
      df_di[n_track,4] <- m_ent
      n_track <- n_track+1
    }
  }
}

```

## Plotting Diversity Scores
```{r}
library(ggplot2)

df_di_cs <- df_di[df_di$Diagnosis == 'CS',]
df_di_szsafd <- df_di[df_di$Diagnosis == 'SZSAFD',]

ggplot(data=df_di, aes(x=Year, y=mult_ent, group=City)) +
  geom_line(data = df_di[df_di$Diagnosis == 'CS',], aes(color = City), linetype='solid', size = 1.1) +
  geom_line(data = df_di[df_di$Diagnosis == 'SZSAFD',], linetype='dashed', aes(color = City), size = 1.1) +
  geom_point(aes(color = City)) +
  ggtitle('Cumulative Diversity Indices\nover Time for HC (solid) and SZSAFD (dashed)')

# szsafd <- ggplot(data=df_di_szsafd, aes(x=Year, y=mult_ent, group=City)) +
#   geom_line(linetype='dashed', aes(color=City)) +
#   ggtitle('Diversity Indices over Time for SZSAFD')


```

## Longitudinal Data Analysis - Framing the question
* Goal: predict diversity index in sample based on:
1) City diversity index - covariate
2) City population (estimate from ACS) - covariate
3) Cumulative Sample Size from that city - covariate
4) Diagnosis designation (variable of interest)

Resource: 

Lalonde, T.L. (2015). Modeling Time-Dependent Covariates in Longitudinal Data Analyses. In: Chen, DG., Wilson, J. (eds) Innovative Statistical Methods for Public Health Data. ICSA Book Series in Statistics. Springer, Cham. https://doi.org/10.1007/978-3-319-18536-1_4

From the graph:
* Can test a diagnosis x time effect within each city 
- See if the responses are parallel or not
* Can see how diversity index profiles vary based on city characteristics

## Testing Diagnosis x time effect
NOTE: no time-dependent covariates involved in model
```{r}
library(nlme)
library(lme4)

df_di$Post10 <- df_di$Year-2010
max_di <- log(8)
df_di$mult_ent_s <- df_di$mult_ent / max_di * 100

# random intercept and slopes model fails to converge
# random intercept model instead
ri_mod <- lme(fixed = mult_ent_s ~ Diagnosis * factor(Post10), random = ~ 1|City,
                  method = "REML", na.action = na.omit, data=df_di)
summary(ri_mod)

ri_mod_ct <- lme(fixed = mult_ent_s ~ Diagnosis * Post10, random = ~ 1|City,
                  method = "REML", na.action = na.omit, data=df_di)
summary(ri_mod_ct)

ri_mod_ml <- lme(fixed = mult_ent_s ~ Diagnosis * factor(Post10), random = ~ 1|City,
                  method = "ML", na.action = na.omit, data=df_di)

ri_mod_ct_ml <- lme(fixed = mult_ent_s ~ Diagnosis * Post10, random = ~ 1|City,
                  method = "ML", na.action = na.omit, data=df_di)

anova(ri_mod_ct_ml, ri_mod_ml)

# no evidence for response profile fitting significantly better than continuous time
summary(ri_mod_ct)
```

Preliminary conclusion:

After accounting for city-level variation at baseline, diversity index increased by 2.30351 percentage points per year (p = 0.0303). Relative to healthy controls, the SZSAFD condition experienced a slightly attenuated increase (-0.81081) in diversity index, though this attenuation is not statistically significant (p = 0.5798).

Moving forward:
- Try to use ethnicity and gender --> by-city differences
-- Just only use races for diversity index, then do follow-up analyses on ethnicity and gender
-- do the race only, then separate analyses on eth and gender
-- then see if you can combine race/eth/gender
-- find out about the 200 missing people

Thoughts for regression:
* What if you used percentage of max diversity score?
