---
title: "56. COGS Linear Regression"
author: "Daniel Zoleikhaeian"
date: "2023-08-23"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# loading dataset and libraries
```{r}
library(car)
library(readxl)
library(ggplot2)
library(GGally)
library(splines)
library(plyr)
library(effects)
df <- read_xlsx('C:/Users/danie/Documents/Joshi Lab Materials/3 Studies Dataset/Dataset Merge/COGS_data.xlsx')

colnames(df)
col_sub <- c('SEX', 'GROUP', 'RACE','AGE','ETHNICITY','EDU_SELF','MMN135_205')

df_sub <- df[, col_sub]
head(df_sub)
```

# Sample linear regression on EDU_SELF

## Checking missing data and assumptions
```{r}
# checking for missing data
sum(is.na(df_sub$EDU_SELF)) # only 1/2477 subjects don't have EDU_SELF
sum(complete.cases(df_sub[, -7])) # 2476/2477 have complete data

# getting rid of that one individual
df_self <- df_sub[complete.cases(df_sub[, -7]), -7]

# pairs plot of data to check normality and correlations
ggpairs(df_self)
# no strong correlations among the predictors
# continuous variables are roughly normally distributed
```

## Checking the levels on the factorized variables
These are ethnicity, race, sex, group

Race encodings:
1 = 'NH'
2 = 'AS'
3 = 'PI'
4 = 'AA'
5 = 'CA'
6 = 'MR'
7 = 'UNK'

```{r}
unique(df_self$ETHNICITY) # 2 or 1
unique(df_self$RACE) # 1 thru 7 
unique(df_self$GROUP) # 1 or 2 --> 1 is control, 2 is SZ
unique(df_self$SEX) # M or F

dict_transform <- function(keys, values, vec) {
  for (i in 1:length(keys)) {
    vec[vec == keys[i]] <- values[i]
  }
  return (vec)
}

# transforming the categorical variables by a dictionary
df_self$ETHNICITY2 <- factor(dict_transform(c(1,2), c('Hispanic', 'NotHispanic'), df_self$ETHNICITY) )
df_self$RACE2 <- factor(dict_transform(1:7, c('NH','AS','PI','AA','CA','MR','UNK'), df_self$RACE))
df_self$GROUP2 <- factor(dict_transform(c(1,2), c('CS', 'SZSAFD'), df_self$GROUP))

# releveling
df_self$ETHNICITY2 <- relevel(df_self$ETHNICITY2, ref = 'NotHispanic')
df_self$RACE2 <- relevel(df_self$RACE2, ref = 'AA')
df_self$GROUP2 <- relevel(df_self$GROUP2, ref = 'CS')
df_self$SEX <- relevel(factor(df_self$SEX), ref = 'F') # females are the reference
```

## Performing the Regression
```{r}
edu_lr <- lm(EDU_SELF ~ SEX + RACE2 + ETHNICITY2 + GROUP2 + AGE, data = df_self)

# checking fit
plot(edu_lr, 1) # errors have mean 0; suspected high variance in middle than in end
plot(edu_lr, 2) # residuals are mostly normally distributed
plot(edu_lr, 4) # a few high-leverage points

# checking for curvature and added complexity
residualPlots(edu_lr) # p = 0.048, some added complexity required, but no evidence for quadratic term on Age (p = 0.77)

## attempt: Diagnosis x Sex
edu_lr2 <- lm(EDU_SELF ~ SEX * GROUP2 + RACE2 + ETHNICITY2 + AGE, data = df_self)

residualPlots(edu_lr2) # p = 0.1842, no added complexity required

# checking non-constant variance
ncvTest(edu_lr2) # p = 0.10, so there is no evidence of ncv

# verifying ncv by looking at residuals
plot(edu_lr2, 1) # fairly uniform, but still a little suspicious


# estimating the weights to address NCV
plot((residuals(edu_lr2))^2 ~ fitted(edu_lr2), data = df_self,
     main = 'Estimating Weights')

## looks like a 2nd degree polynomial would estimate the variance
res_mod <- lm(((residuals(edu_lr2))^2) ~ poly(fitted(edu_lr2), degree = 3, raw = TRUE)) # one extra degree to be safe
lines(fitted(edu_lr2), predict(res_mod), col = 'blue')
summary(res_mod)
wts <- 1/(predict(res_mod))

wls_edu_lr <- lm(EDU_SELF ~ SEX * GROUP2 + RACE2 + ETHNICITY2 + AGE, data = df_self, weights = wts)
ncvTest(wls_edu_lr) # p = 0.995 non-constant variance addressed

# checking results
summary(wls_edu_lr)

# comparing results to before
summary(edu_lr2) # not much changed
```

Result Highlights:
* All races except NH and PI had significantly higher self edu than AA
* Hispanics had significantly lower self edu compared to non-Hispanics
* SZSAFD group had significantly lower self edu compared to CS
* Male SZSAFD had significantly higher self edu compared to female CS

# Same analysis for MMN135-205
```{r}
# counting up the missing data
sum(is.na(df_sub$MMN135_205)) # 849/2477 subjects don't have MMN135_205
nrow(df_sub)

# just getting rid of the education column
df_MMN <- df_sub[,-6]
colnames(df_MMN)

# omitting missing data for preliminary analysis
df_MMN <- df_MMN[complete.cases(df_MMN),] # 1628, as expected 

# transforming the categorical variables by a dictionary
df_MMN$ETHNICITY2 <- factor(dict_transform(c(1,2), c('Hispanic', 'NotHispanic'), df_MMN$ETHNICITY) )
df_MMN$RACE2 <- factor(dict_transform(1:7, c('NH','AS','PI','AA','CA','MR','UNK'), df_MMN$RACE))
df_MMN$GROUP2 <- factor(dict_transform(c(1,2), c('CS', 'SZSAFD'), df_MMN$GROUP))

# releveling
df_MMN$ETHNICITY2 <- relevel(df_MMN$ETHNICITY2, ref = 'NotHispanic')
df_MMN$RACE2 <- relevel(df_MMN$RACE2, ref = 'AA')
df_MMN$GROUP2 <- relevel(df_MMN$GROUP2, ref = 'CS')

# checking distributions

colnames(df_MMN)
ggpairs(df_MMN[, c('AGE','MMN135_205')])# Age is slightly bimodal, MMN is a little left skewed, but should be fine

```

## Performing the regression
```{r}
mmn_lr <- lm(MMN135_205 ~ SEX + RACE2 + ETHNICITY2 + GROUP2 + AGE, data = df_MMN)

# checking fit
plot(mmn_lr, 1) # errors have mean; variance looks mostly uniform, except at beginning; no obvious curvature
plot(mmn_lr, 2) # residuals are mostly normally distributed but deviate a lot at the tails
plot(mmn_lr, 4) # very few influential points

# checking for curvature
residualPlots(mmn_lr) # p = 0.002 suggests adding higher complexity terms (interactions, but no specific addition of quadratic term)

residualPlot(mmn_lr) # curvature for low and high fitted values

avPlots(mmn_lr)

mmn_lr_itx <- lm(MMN135_205 ~ AGE*RACE2 + ETHNICITY2 + GROUP2 + SEX, data = df_MMN)

plot(Effect(c("RACE2", "AGE"), mmn_lr_itx, default.levels=100),
     rug=FALSE, grid=TRUE, multiline=TRUE, x.var = "AGE")

residualPlots(mmn_lr_itx)

mmn_lr_itx2 <- lm(MMN135_205 ~ RACE2 + ETHNICITY2 + SEX*AGE + GROUP2, data = df_MMN)

plot(Effect(c("SEX", "AGE"), mmn_lr_itx2, default.levels=100),
     rug=FALSE, grid=TRUE, multiline=TRUE, x.var = "AGE")

residualPlots(mmn_lr_itx2)

mmn_lr_itx3 <- lm(MMN135_205 ~ RACE2 + ETHNICITY2 + SEX + AGE * GROUP2, data = df_MMN)

plot(Effect(c("GROUP2", "AGE"), mmn_lr_itx2, default.levels=100),
     rug=FALSE, grid=TRUE, multiline=TRUE, x.var = "AGE") # lines are parallel, so no group x age effect

mmn_lr_itxc <- lm(MMN135_205 ~ RACE2 + ETHNICITY2 + SEX + AGE + GROUP2 + SEX:AGE + RACE2:AGE, data = df_MMN)

residualPlots(mmn_lr_itxc) # p = .247; complexity resolved by adding 2nd degree interactions

# checking non-constant variance
ncvTest(mmn_lr_itxc) # p << 0.05, so there is evidence of ncv

# estimating the weights to address NCV
plot((residuals(mmn_lr_itxc))^2 ~ fitted(mmn_lr_itxc), data = df_MMN,
     main = 'Estimating Weights') # high variance towards the early-middle

## looks like a 2nd degree polynomial would estimate the weights
res_mod_mn <- lm(((residuals(mmn_lr_itxc))^2) ~ poly(fitted(mmn_lr_itxc), degree = 2, raw = TRUE)) # one extra dergee for good luck
lines(fitted(mmn_lr_itxc), predict(res_mod_mn), col = 'blue')
summary(res_mod_mn)
wts <- 1/(predict(res_mod_mn))

wls_mmn <- lm(MMN135_205 ~ RACE2 + ETHNICITY2 + SEX + AGE + GROUP2 + SEX:AGE + RACE2:AGE, data = df_MMN, weights = wts)
ncvTest(wls_mmn) # p = 0.908 non-constant variance addressed

# checking results
summary(wls_mmn)
```

Main findings:
* Race modulates the slope on Age; varying levels of significantly increased MMN
* Higher age = higher MMN
* Higher overall MMN for SZSASFD and Males


## Trying maximal model
```{r}
## attempt: AGE x SEX x Diagnosis x RACE
mmn_lr4 <- lm(MMN135_205 ~ AGE * SEX * GROUP2 * RACE2 * ETHNICITY2, data = df_MMN)
summary(mmn_lr4)

residualPlots(mmn_lr4) # p = 0.2974, so no added complexity required

# ISSUE: this model is not very interpretable, so use the 2nd degree interaction model

# checking non-constant variance
ncvTest(mmn_lr4) # p << 0.05, so there is evidence of ncv

# estimating the weights to address NCV
plot((residuals(mmn_lr4))^2 ~ fitted(mmn_lr4), data = df_MMN,
     main = 'Estimating Weights') # high variance towards the early-middle

## looks like a 2nd degree polynomial would estimate the weights
res_mod_mn <- lm(((residuals(mmn_lr4))^2) ~ poly(fitted(mmn_lr4), degree = 3, raw = TRUE)) # one extra dergee for good luck
lines(fitted(mmn_lr4), predict(res_mod_mn), col = 'blue')
summary(res_mod_mn)
wts <- 1/(predict(res_mod_mn))

wls_mmn2 <- lm(MMN135_205 ~ AGE * SEX * GROUP2 * RACE2 +  ETHNICITY2, data = df_MMN, weights = wts)
ncvTest(wls_mmn2) # p = 0.934 non-constant variance addressed

# checking results
summary(wls_mmn2)
```

## Comparing Maximal Model to 2nd degree interaction model
```{r}
# 2nd degree interaction model
summary(wls_mmn)$adj.r.squared # 0.2938182

# maximal model
summary(wls_mmn2)$adj.r.squared # 0.2949332
```

# Checking if the MMN data is not Missing at Random
```{r}
df_MMN2 <- df_sub[,-6]
colnames(df_MMN2)

df_MMN2$cmpl <- ifelse(is.na(df_MMN2$MMN135_205), 1, 0) # 1 = missing, 0 = complete

# applying logistic regression: goal would be to see no trends
mar_check <- glm(cmpl ~ SEX + factor(RACE) + factor(ETHNICITY) + factor(GROUP) + AGE, data = df_MMN2, family = 'binomial')

# checking the fit
library(glmtoolbox)
hltest(mar_check)
# suggests complexity is not sufficient (p = 0.12 < 0.20 cutoff)

# attempt 2: using same formula lm
mar_check2 <- glm(cmpl ~ factor(RACE) + factor(ETHNICITY) + factor(SEX) + AGE + factor(GROUP) + factor(SEX):AGE + factor(RACE):AGE, data = df_MMN2, family = 'binomial')
hltest(mar_check2) # p = 0.204; no additional complexity required

summary(mar_check2)
library(rje)
exp(coef(mar_check2))
exp(confint(mar_check2))

Anova(mar_check2)

# last sanity check
df_MMN_inc <- df_MMN2[is.na(df_MMN2$MMN135_205),]
cts <- plyr::count(df_MMN_inc, 'GROUP')
cts$prop <- cts$freq / sum(cts$freq)
cts

prop.test(310, 310+539, p = 0.5) # p = 5.079e-15; not a 50-50 split between CS and SZSAFD among those who are missing MMN
```

Main findings:
* Being in the SZ group has higher odds (1.47; 95% CI: 1.22 to 1.77) of not having an MMN recorded. Controlling for Age, Race, Ethnicity, and Sex, the odds of having a missing MMN are 50% higher for SZ patients. 
* Moderate suggestion that older males have lower odds of dropout than older females
* No significant increases in the odds of missing MMN for other variables

Conclusion:
* Data is not missing at random
* SZSAFD subjects have significantly higher chance of having a missing MMN