---
title: "58. Added Variable Analysis COGS"
author: "Daniel Zoleikhaeian"
date: "2023-09-06"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# loading dataset and libraries
```{r}
library(car)
library(readxl)
library(ggplot2)
library(GGally)
library(splines)
library(plyr)
library(effects)
df <- read_xlsx('C:/Users/danie/Documents/Joshi Lab Materials/3 Studies Dataset/Dataset Merge/COGS_data.xlsx')

colnames(df)
col_sub <- c('SEX', 'GROUP', 'RACE','AGE','ETHNICITY','MMN135_205')

df_sub <- df[, col_sub]
head(df_sub)

colnames(df_sub)

# omitting missing data
df_sub <- df_sub[complete.cases(df_sub), ] # 1628, as expected

# transforming the categorical variables by a dictionary
dict_transform <- function(keys, values, vec) {
  for (i in 1:length(keys)) {
    vec[vec == keys[i]] <- values[i]
  }
  return (vec)
}

df_sub$ETHNICITY2 <- factor(dict_transform(c(1,2), c('Yes', 'No'), df_sub$ETHNICITY) )
df_sub$RACE2 <- factor(dict_transform(1:7, c('AE','AS','NH','AA','CA','MR','UNK'), df_sub$RACE))
df_sub$GROUP2 <- factor(dict_transform(c(1,2), c('CS', 'SZSAFD'), df_sub$GROUP))

# releveling
df_sub$ETHNICITY2 <- relevel(df_sub$ETHNICITY2, ref = 'No')
df_sub$RACE2 <- relevel(df_sub$RACE2, ref = 'AA')
df_sub$GROUP2 <- relevel(df_sub$GROUP2, ref = 'CS')

# subsetting by group
cs <- df_sub[df_sub$GROUP == 1, ]
sz <- df_sub[df_sub$GROUP == 2, ]
```

# Added Variable Analysis
```{r}
mmn_cs <- lm(MMN135_205 ~ AGE, data = cs)
summary(mmn_cs) # R2 = 0.1332
res_cs <- resid(mmn_cs)
dem_cs <-lm(res_cs ~ SEX + RACE2 + ETHNICITY2, data = cs)
summary(dem_cs) # R2 = 0.06298

mmn_sz <- lm(MMN135_205 ~ AGE, data = sz)
summary(mmn_sz) # R2 = 0.1221
res_sz <- resid(mmn_sz)
dem_sz <-lm(res_sz ~ SEX + RACE2 + ETHNICITY2, data = sz)
summary(dem_sz) # R2 = 0.02636
```

Preliminary conclusion:
* For SZ, demographic variables are less explanatory of the variance explained by AGE compared to in CS

# Separate Site Analysis

## Prepping the overall dataframe
```{r}
df$ETHNICITY2 <- factor(dict_transform(c(1,2), c('Yes', 'No'), df$ETHNICITY) )
df$RACE2 <- factor(dict_transform(1:7, c('AE','AS','NH','AA','CA','MR','UNK'), df$RACE))
df$GROUP2 <- factor(dict_transform(c(1,2), c('CS', 'SZSAFD'), df$GROUP))

# releveling
df$ETHNICITY2 <- relevel(df$ETHNICITY2, ref = 'No')
df$RACE2 <- relevel(df$RACE2, ref = 'AA')
df$GROUP2 <- relevel(df$GROUP2, ref = 'CS')
```


```{r}
sites <- unique(df$SITEID) # 5 different sites, coded 1 through 5
sites[6] <- 6 # add a "6th site" for the pooled dataset

lm_list <- vector(mode = 'list', length = 6)
names(lm_list) <- c('UCLA', 'UCSD', 'Sinai', 'Penn', 'UW', 'Pool')


for (i in sites) {
  
  list_in <- vector(mode = 'list', length = 4)
  names(list_in) <- c('CS', 'SZ', 'CS aov','SZ aov')
  
  if (i != 6) {
    
    mod_cs <- lm(MMN135_205 ~ AGE + SEX + RACE2 + ETHNICITY2, data = df[df$SITEID == i & df$GROUP == 1, ]) 
    cs_res <- as.data.frame(summary(mod_cs)$coefficients)
    cs_res$sig_code <- ifelse(cs_res$`Pr(>|t|)` > 0.05, '-', '*')
    
    mod_sz <- lm(MMN135_205 ~ AGE + SEX + RACE2 + ETHNICITY2, data = df[df$SITEID == i & df$GROUP == 2, ])
    sz_res <- as.data.frame(summary(mod_sz)$coefficients)
    sz_res$sig_code <- ifelse(sz_res$`Pr(>|t|)` > 0.05, '-', '*')
    
    list_in[[1]] <- cs_res
    list_in[[2]] <- sz_res
    list_in[[3]] <- Anova(mod_cs)
    list_in[[4]] <- Anova(mod_sz)
    
    lm_list[[i]] <- list_in
    
    
    
  } else {
    
    mod_cs <- lm(MMN135_205 ~ AGE + SEX + RACE2 + ETHNICITY2, data = df[df$GROUP == 1, ]) 
    cs_res <- as.data.frame(summary(mod_cs)$coefficients)
    cs_res$sig_code <- ifelse(cs_res$`Pr(>|t|)` > 0.05, '-', '*')
    
    mod_sz <- lm(MMN135_205 ~ AGE + SEX + RACE2 + ETHNICITY2, data = df[df$GROUP == 2, ])
    sz_res <- as.data.frame(summary(mod_sz)$coefficients)
    sz_res$sig_code <- ifelse(sz_res$`Pr(>|t|)` > 0.05, '-', '*')
    
   list_in[[1]] <- cs_res
   list_in[[2]] <- sz_res
   list_in[[3]] <- Anova(mod_cs)
   list_in[[4]] <- Anova(mod_sz)
   
   lm_list[[i]] <- list_in
    
  }
}

lm_list

```

* Different sites had different findings

* AGE, SEX, and RACE are significant in the pool for both diagnoses
- SEX and RACE not significant in any of the sites for SZ
- Proposed explanation: these SEX and RACE are weaker predictors of MMN, so larger sample size is needed to detect a significant effect

* AGE not significant in:
- Sinai CS
* SEX not significant in:
- UW SZ and CS
- Penn SZ
- Sinai SZ and CS
- SD CS and SZ
- LA SZ
* RACE not significant in:
- LA CS and SZ
- SD SZ
- Sinai CS and SZ
- Penn SZ
- UW SZ

## Checking the sample sizes
```{r}
plyr::count(df[df$GROUP==1, ], 'SITEID') # Roughly equal sample sizes among CS
plyr::count(df[df$GROUP==2, ], 'SITEID') # UCLA dominates among SZ
```